---
title: "Take-home-Exercise 2"
author: "Li Yuquan"
description: ""
date: "October 5, 2025"
date-modified: "last-modified"
format: html
editor: visual
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# 1 Overview

Urban mobility is a crucial component of sustainable city development, influencing not only transportation efficiency but also environmental impact and quality of life. In rapidly urbanizing cities like Singapore, public bus services play a central role in connecting communities and supporting daily commuting activities. With the increasing digitalization of transportation infrastructure — including smart cards, GPS systems, and real-time passenger tracking — large volumes of spatio-temporal data are now available to analyze movement patterns more precisely than ever before.

However, much of this data remains underutilized, as conventional GIS-based approaches often focus on simple mapping and visualization rather than rigorous spatial statistics. To address this gap, this study aims to apply advanced geospatial analytical techniques to uncover deeper insights into Singapore’s bus passenger mobility. Specifically, it focuses on identifying localized clustering patterns and temporal dynamics of travel demand, using Local Measures of Spatial Autocorrelation (LMSA) and Emerging Hot Spot Analysis (EHSA).

## 1.1 Objective

The objectives of this project are therefore threefold:

1.  To examine how bus passenger trips vary spatially across different neighborhoods and time periods, revealing localized areas of high and low travel demand.

2.  To identify statistically significant hot spots, cold spots, and outliers in passenger volumes using LMSA methods such as Local Moran’s I, Local Geary’s C, and Getis-Ord Gi\*.

3.  To detect evolving mobility patterns through EHSA, categorizing temporal changes in hot and cold spots (e.g., new, intensifying, diminishing).

By integrating spatial and temporal perspectives, the study seeks to enhance understanding of urban mobility behavior and provide data-driven insights for smarter and more adaptive transport planning in Singapore.

## 1.2 The Task

The first component, Geospatial Data Science, involves the construction of a spatial analytical framework by deriving a hexagonal grid system with a 375-meter resolution to represent Traffic Analysis Zones (TAZs) across Singapore. Using passenger trip records provided by the Land Transport Authority (LTA) DataMall, the total number of trips originating from each hexagon is computed for three key time intervals: weekday morning peak (6 a.m.–9 a.m.), weekday afternoon peak (5 p.m.–8 p.m.), and weekend/holiday (11 a.m.–8 p.m.). The resulting datasets are visualized through thematic maps to explore the spatial distribution of passenger flows. These visualizations provide preliminary insights into how bus mobility varies between neighborhoods and across different periods of the week.

The second component, Local Measures of Spatial Autocorrelation (LMSA), applies local spatial statistics—including Local Moran’s I, Local Geary’s C, and Getis-Ord Gi\*—to the hexagon-level passenger data. Unlike global indices that summarize overall spatial relationships, these local measures allow for the identification of statistically significant clusters of high or low values (hotspots and coldspots) and spatial outliers (locations that differ sharply from their surroundings). By visualizing only statistically significant areas (p \< 0.05), this step reveals the geographic concentration and intensity of commuting demand within the city.

The third component, Emerging Hot Spot Analysis (EHSA), extends the spatial analysis into the temporal domain. Using a space-time cube representation of passenger trip data, EHSA combines the local Gi\* statistic with the Mann–Kendall trend test to identify not only where hotspots and coldspots occur, but also how they evolve over time. The analysis classifies areas into categories such as new, intensifying, diminishing, or sporadic hot and cold spots, offering a dynamic view of how commuting patterns shift between weekdays and weekends.

## 1.3 R Packages

```{r}
pacman::p_load( skimr, knitr, DT, performance, reshape2, ggpubr, ggplot2, sf, sp, sfdep, tmap, stplanr, tidyverse,plotly, h3jsr)
```

## 1.4 The Data

# 2 Data Import and Wrangling

## 2.1 Import Geospatial data

### 2.1.1 mpsz data

```{r}
mpsz_sf <- st_read("data/geospatial/MasterPlan2019SubzoneBoundaryNoSeaKML.kml") %>%
  st_zm(drop = TRUE, what = "ZM") %>% st_transform(crs = 3414)
mpsz <- mpsz_sf
```

```{r}
summary(mpsz_sf)
```

```{r}
glimpse(mpsz_sf)
```

### 2.1.2 bus stop data

```{r}
BusStop = st_read(dsn = "data/LTADataMall/",
                  layer = "BusStop") %>%
  st_transform(crs = 3414)
```

```{r}
summary(BusStop)
```

```{r}
glimpse(BusStop)
```

## 2.2 Import Apstial data

### 2.2.1 Passenger Volume by Origin Destination Bus Stops data

```{r}
odbus <- read_csv("data/LTADataMall/origin_destination_bus_202508.csv")
```

```{r}
summary(odbus)
```

```{r}
glimpse(odbus)
```

## 2.3 Data Prepare

### 2.3.1 Remove duplicate bus stops

```{r}
BusStop <- BusStop %>% distinct(BUS_STOP_N, .keep_all = TRUE)


BusStop %>% 
  st_drop_geometry() %>% 
  group_by(BUS_STOP_N) %>%
  filter(n()>1) %>%
  ungroup() %>% 
  arrange(BUS_STOP_N)
```

### 2.3.2 Clip to Singapore boundary

```{r}
BusStop_in_SG <- st_join(
  BusStop, mpsz,
  join = st_within,
  left = FALSE
)
```

### Deriving Analytical Hexagon

```{r}
hexagon <- st_make_grid(mpsz,
                        cellsize = 750,
                        what = "polygon",
                        square = FALSE) %>%
  st_sf()
```

### Fitering hexagons with bus stops

```{r}
hexagon$busstop_count =
  lengths(st_intersects(
    hexagon, BusStop_in_SG))

hexagon <- filter(hexagon, busstop_count > 0)
```

###Assigning ids to each hexagon

```{r}
hexagon <- hexagon %>% 
  select(, -busstop_count)

hexagon$HEX_ID <- sprintf(
  "H%04d", seq_len(
    nrow(hexagon))) %>% 
  as.factor()
```

## Preparing Trip Generation Data

### Cleaning the data

```{r}
trips <- odbus %>%
  select(c(ORIGIN_PT_CODE, DAY_TYPE, TIME_PER_HOUR, TOTAL_TRIPS)) %>%
  rename(BUS_STOP_N = ORIGIN_PT_CODE) %>%
  rename(HOUR_OF_DAY = TIME_PER_HOUR) %>%
  rename(TRIPS = TOTAL_TRIPS)

trips$BUS_STOP_N <- as.factor(trips$BUS_STOP_N)

kable(head(trips))
```

### Combine

```{r}
bs_hex <- st_intersection(
  BusStop, hexagon) %>%
  st_drop_geometry() %>%
  select(c(BUS_STOP_N, HEX_ID))

kable(head(bs_hex))
```

### Adding HEX_ID into bus trips data

```{r}
trips <- inner_join(trips, bs_hex)
kable(head(trips))
```

### Aggregating TRIPS based on HEX_ID

```{r}
trips <- trips %>%
  group_by(
    HEX_ID,
    DAY_TYPE,
    HOUR_OF_DAY) %>%
  summarise(TRIPS = sum(TRIPS))
```

## Visualize

```{r}
tm_shape(mpsz) +
  tm_polygons() +
tm_shape(BusStop_in_SG) +
  tm_dots()
```

```{r}
hex_active <- hexagon %>%
  mutate(busstop_count = lengths(st_intersects(., BusStop_in_SG))) %>%
  filter(busstop_count > 0)

bs_in_active_hex <- st_join(BusStop_in_SG, hex_active, join = st_within, left = FALSE)

tm_shape(mpsz) +
  tm_polygons(col = "#d7f2d0", border.col = "white") +
tm_shape(hex_active) +
  tm_borders(col = "grey30", lwd = 0.6) +
tm_shape(bs_in_active_hex) +
  tm_dots(size = 0.02, col = "black", alpha = 0.7) +
tm_layout(main.title = "Hexagons & Bus Stops",
          legend.show = FALSE, frame = FALSE)
```

```{r}
glimpse(hex_active)
```

```{r}
plot(hex_active['busstop_count'],
     main = 'Number of bus stops per hexagon')

```


```{r}
glimpse(hex_active)
```
# LISA


## start



```{r}
# 1) choose a time slice (example: WEEKDAY, 8am)
slice <- trips %>%
  filter(DAY_TYPE == "WEEKDAY", HOUR_OF_DAY == 8) %>%
  select(HEX_ID, TRIPS)

# 2) join to hex polygons (use hex_active if that’s your filtered layer)
hex_slice <- hex_active %>%
  left_join(slice, by = "HEX_ID") %>%
  mutate(TRIPS = ifelse(is.na(TRIPS), 0, TRIPS))  # fill zeros if no trips at that hour

```

## 1 Computing Contiguity Spatial Weights

```{r}

wm_q <- hex_slice %>%
  mutate(nb = st_contiguity(geometry),
         wt = st_weights(
           nb, style = "W", allow_zero = TRUE),
         .before = 1) 
```

```{r}
summary(wm_q$nb)
```
## 2 Computing local Moran’s I

```{r}
glimpse(wm_q)
```



```{r}
set.seed(1234)
```

```{r}
options(sfdep.p_adjust_significance_breaks = NULL)
options(sfdep.p_adjust_significance_labels = NULL)

# 可选：看一眼现在的配置，确认无误
getOption("sfdep.p_adjust_significance_breaks")
getOption("sfdep.p_adjust_significance_labels")
```

```{r}
lisa <- wm_q %>% 
  mutate(local_moran = local_moran(
    TRIPS, nb, wt, nsim = 99),
         .before = 1) %>%
  unnest(local_moran)
```




