---
title: "In class Ex 9"
author: "Li Yuquan"
description: ""
date: "November 8, 2025"
date-modified: "November 8, 2025"
format: html
editor: visual
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

```{r}
pacman::p_load(tmap, SpatialAcc, sf, 
               ggstatsplot, reshape2,
               tidyverse)
```

```{r}
mpsz <- st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_NO_SEA_PL")
hexagons <- st_read(dsn = "data/geospatial", layer = "hexagons") 
eldercare <- st_read(dsn = "data/geospatial", layer = "ELDERCARE") 

mpsz <- st_transform(mpsz, 3414)
eldercare <- st_transform(eldercare, 3414)
hexagons <- st_transform(hexagons, 3414)

```

```{r}
eldercare <- eldercare %>%
  select(fid, ADDRESSPOS) %>%
  mutate(capacity = 100)

hexagons <- hexagons %>%
  select(fid) %>%
  mutate(demand = 100)
```

```{r}
ODMatrix <- read_csv("data/aspatial/OD_Matrix.csv")
distmat <- ODMatrix %>%
  select(origin_id, destination_id, total_cost) %>%
  spread(destination_id, total_cost) %>%
  select(-origin_id)
distmat_km <- as.matrix(distmat / 1000)
```

```{r}
powers <- c(0.5, 1, 1.5, 2, 2.5)

results <- list()
for (p in powers) {
  acc <- data.frame(ac(hexagons$demand,
                       eldercare$capacity,
                       distmat_km,
                       power = p,
                       family = "Hansen"))
  colnames(acc) <- "accHansen"
  acc <- as_tibble(acc)
  hex_res <- bind_cols(hexagons, acc)
  hex_res$power <- p
  results[[as.character(p)]] <- hex_res
}

hexagon_Hansen_all <- bind_rows(results)
```

```{r}
mapex <- st_bbox(hexagons)
```

This figure shows the Hansen accessibility to eldercare facilities in Singapore under different power values. When the power is low , the effect of distance is weak, and accessibility values are more evenly distributed across the island. As the power increases (1.5 - 2.5), the impact of distance becomes stronger—areas closer to eldercare facilities show much higher accessibility, while distant areas experience a sharp decline. Higher power values therefore highlight stronger spatial inequality in service accessibility.

```{r, fig.height= 10, fig.width= 10}
tm_shape(hexagon_Hansen_all, bbox = mapex) +
  tm_facets(by = "power", ncol = 3) +
  tm_polygons(fill = "accHansen",
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 10,
                values = "brewer.blues")) +
  tm_shape(eldercare) +
  tm_dots(size = 0.3, fill = "red", col = "black") +
  tm_title("Accessibility to eldercare (Hansen, different power values)") +
  tm_layout(frame = TRUE) +
  tm_compass(type = "8star", size = 2) +
  tm_scalebar() +
  tm_grid(alpha = 0.2)
```
