---
title: "Take-home Exercise 1"
author: "Li Yuquan"
description: ""
date: "September 15, 2025"
date-modified: "last-modified"
format: html
editor: visual
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# 1 Overview

## 1.1 Objectives

The location choices of businesses play a pivotal role in shaping urban development, influencing economic growth, job creation, and the provision of resources and services. Strategic siting not only creates competitive advantages and supports efficient supply chains, but also contributes to a city’s identity through specialized districts and accessible infrastructure.

Studying the spatio-temporal patterns of new businesses is therefore essential, as it highlights how economic activity and urban forms evolve over time. Such analysis enables urban planners to anticipate growth, identify clusters and trends, improve resource allocation, and design more sustainable and livable cities.

In this exercise, the focus is on new business entities established in Singapore during the first six months of 2025. The objectives are:

• To analyze the “where” and “when” of new business establishments, uncovering spatial clusters, temporal trends, and development hotspots.

• To explore complex interaction patterns beyond simple density measures, such as the clustering of related firms, path-dependency of business locations, and structural changes in commercial areas.

## 1.2 The Task

1.  Download ACRA (Accounting and Corporate Regulatory Authority) Information on Corporate Entities datasets from data.gov.sg.

2.  Select at least three business types by referring to [2-digit or 3-digit Singapore Standard Industrial Classification (SSIS) 2020](https://classification.codes/classifications/industry/ssic-singapore). Prepare and tidy the data by applying appropriate data wrangling method.

3.  Extract the selected business entities registered between 1st January 2024 to 30th June 2025.

4.  Geocode the selected records and convert them into simple point features.

5.  Compute spatial and spatio-temporal KDE of the selected business types.

6.  Perform second order spatial and spatio-temporal analysis on the selected business types.

7.  Describe and discuss the analysis results obtained from Step 5 and 6.

8.  For each geovisualisation prepared, write a short report of not more than 150 words describing the spatio-temporal patterns revealed by the geovisualisation and statistical graphics.

9.  With reference to the analysis results above, discuss how these findings can be used to support urban land use planning and management. (Not more than 500 words)

# 1.3 R packages

| **Package** | **Description** |
|------------------------------------|------------------------------------|
| sf | Simple Features for handling and analyzing spatial vector data. |
| spNetwork | Network-based spatial analysis, e.g. kernel density along road networks. |
| spatstat | Spatial point pattern analysis (ppp, owin, G/F/K/L functions, KDE). |
| tmap | Thematic maps creation (static and interactive). |
| tidyverse | Data science suite (dplyr, ggplot2, readr, purrr, etc.). |
| httr | Tools for working with HTTP to access web APIs. |
| rvest | Web scraping and HTML parsing. |
| terra | Spatial raster data handling and analysis. |
| lwgeom | Geometry operations, fixing invalid shapes (wraps GEOS). |

```{r}
pacman::p_load(sf, spNetwork, spatstat,tmap, tidyverse, httr, rvest, terra, lwgeom)
```

# 1.4 The Data

| **Dataset Name** | **Description** | **Format** | **Source** |
|------------------|------------------|------------------|------------------|
| ACRA Information on Corporate Entities ('A-Z-Others') | Business registration information (clinics, massage, fitness filtered by SSIC) | CSV | data.gov.sg |
| OneMap Geocoding Results | Geocoded coordinates (X, Y) for postal codes | API | [OneMap API (Singapore Land Authority)](https://onemap.gov.sg/api/common/elastic/search) |
| MasterPlan2019SubzoneBoundaryNoSeaKML | Planning region and subzone polygon boundaries | KML | URA (Urban Redevelopment Authority) |
| Singapore Standard Industrial Classification (SSIS) 2020 | SSIS Code Matching Table | Excel | [2-digit or 3-digit Singapore Standard Industrial Classification (SSIS) 2020](https://classification.codes/classifications/industry/ssic-singapore) |

# 4 Data Import and Wrangling

## 4.1 Importing ACRA data

Import ACRA data to R studio, combin all 27 file togeter and save it as rds file.

```{r}

folder_path <- "data/aspatial"
file_list <- list.files(path = folder_path,
                        pattern = "^ACRA*.*\\.csv$",
                        full.names = TRUE)

acra_data <- file_list %>%
  map_dfr(read_csv)
```

```{r}
write_rds(acra_data,
          "data/rds/acra_data.rds")
```

```{r}
summary(acra_data) 
```

```{r}
glimpse(acra_data)
```

## 4.2 Importing MasterPlan2019SubzoneBoundaryNoSea data

The Master Plan 2019 Subzone Boundary (KML) file was imported, re-projected to Singapore’s SVY21 coordinate system (EPSG:3414), and saved as an RDS file for later mapping and analysis.

```{r}
mpsz <- st_read("data/MasterPlan2019SubzoneBoundaryNoSeaKML.kml")
mpsz <- mpsz %>% st_transform(crs=3414)
write_rds(mpsz, "data/rds/mpsz.rds")
```

```{r}
summary(mpsz) 
```

```{r}
glimpse(mpsz)
```

## 4.3 Tidying ACRA data

The ACRA dataset was filtered to keep only four SSIC codes representing three industry types: Clinics (86201, 86202), Massage services (96024), and Fitness centres (93111). The data was cleaned (date fields, year/month extraction, postal code formatting) and then split into three subsets: biz_clinic, biz_Massage, and biz_fitness.

| **Industry Type** | **SSIC Code(s)** | **Description** |
|------------------|-----------------|-------------------------------------|
| Clinics | 86201, 86202 | Clinics and other general medical services (Western & Non-Western) |
| Massage Services | 96024 | Massage parlours and foot reflexology |
| Fitness Centres | 93111 | Fitness centres and gymnasiums |

```{r}
ssic_vec <- c("86201", "86202", "96024", "93111")

acra_core <- acra_data %>%
  select(1:24) %>% 
  mutate(
    primary_ssic_code = as.character(primary_ssic_code),
    date  = as.Date(registration_incorporation_date),
    YEAR  = year(date),
    MONTH_NUM  = month(date),
    MONTH_ABBR = month(date, label = TRUE, abbr = TRUE)
  ) %>%
  filter(
    date >= as.Date("2024-01-01"),
    date <= as.Date("2025-06-30"),
    primary_ssic_code %in% ssic_vec
  ) %>%
  mutate(
    postal_code = str_pad(postal_code, width = 6, side = "left", pad = "0")
  )


biz_clinic  <- acra_core %>% filter(primary_ssic_code %in% c("86201", "86202"))
biz_Massage  <- acra_core %>% filter(primary_ssic_code %in% c("96024"))
biz_fitness <- acra_core %>% filter(primary_ssic_code == "93111")
```

## 4.4 Geocoding

The postal codes from the three business subsets (clinics, massage, fitness centres) were extracted and cleaned to remove duplicates and missing values. These unique postal codes were then sent to the OneMap API for geocoding. For each code, the API returned geographic coordinates (X, Y) and address details if found. The results were stored in a found dataset, while unmatched postal codes were recorded separately in a not_found dataset.

```{r}
postcodes <- c(biz_clinic$postal_code,
               biz_Massage$postal_code,
               biz_fitness$postal_code) %>%
  unique() %>%
  na.omit()

url <- "https://onemap.gov.sg/api/common/elastic/search"

found     <- data.frame()
not_found <- data.frame(postcode = character())

for (pc in postcodes) {
  query <- list(
    searchVal = pc,
    returnGeom = "Y",
    getAddrDetails = "Y",
    pageNum = "1"
  )
  res  <- GET(url, query = query)
  json <- content(res)
  
  if (!is.null(json$found) && json$found != 0) {
    df <- as.data.frame(json$results, stringsAsFactors = FALSE)
    df$input_postcode <- pc
    found <- bind_rows(found, df)
  } else {
    not_found <- bind_rows(not_found, data.frame(postcode = pc))
  }
 
}
```

## 4.5 Tidying the geocoded data
The geocoding results were cleaned by selecting only the first 10 relevant fields from the found dataset to keep essential location attributes.

```{r}
found <- found %>%
  select(1:10)
```

## 4.6 Appending the location information

The cleaned geocoded results were joined back to the three business datasets (biz_clinic, biz_Massage, biz_fitness) using postal code as the key, enriching each business record with its spatial coordinates and address details. The updated datasets were then saved as RDS files for later use.

```{r}
biz_clinic  <- biz_clinic  %>% left_join(found, by = c("postal_code" = "POSTAL"))
biz_Massage  <- biz_Massage  %>% left_join(found, by = c("postal_code" = "POSTAL"))
biz_fitness <- biz_fitness %>% left_join(found, by = c("postal_code" = "POSTAL"))
```

```{r}
write_rds(biz_clinic, "data/rds/biz_clinic.rds")
write_rds(biz_Massage, "data/rds/Massage.rds")
write_rds(biz_fitness, "data/rds/biz_fitness.rds")
```

### 4.6.1 Glimpse for 3 industry

By performing the Glimpse operation on the data sets of the three industries, we can determine the amount of data contained in each one by looking at the number of rows.

::: panel-tabset
# Clinics

```{r}
glimpse(biz_clinic_sf)
```

# Massage

```{r}
glimpse(biz_Massage_sf)
```

# Fitness Centres

```{r}
glimpse(biz_fitness_sf)
```
:::
## 4.7 Converting into SF data frame

Each business dataset was transformed into an sf (simple features) object (biz_clinic_sf, biz_Massage_sf, biz_fitness_sf), keeping only records with valid coordinates (X, Y). The coordinate system was set to EPSG:3414 (SVY21 / Singapore TM), enabling spatial mapping and analysis.

```{r}
biz_clinic_sf <- st_as_sf(
  biz_clinic %>% filter(!is.na(X), !is.na(Y)),
  coords = c("X","Y"), crs = 3414
)

biz_Massage_sf <- st_as_sf(
  biz_Massage %>% filter(!is.na(X), !is.na(Y)),
  coords = c("X","Y"), crs = 3414
)

biz_fitness_sf <- st_as_sf(
  biz_fitness %>% filter(!is.na(X), !is.na(Y)),
  coords = c("X","Y"), crs = 3414
)
```

## 4.8 Extracting Attributes from KML Description

The imported Master Plan 2019 Subzone Boundary (mpsz) dataset contained key attributes (e.g., region, planning area, subzone name, subzone code) embedded inside the Description field of the KML file. A custom function extract_kml_field() was created to parse the HTML structure and extract these values. The extracted fields (REGION_N, PLN_AREA_N, SUBZONE_N, SUBZONE_C) were then added as new columns to the dataset. Finally, redundant fields (Name, Description) were removed, and the geometry column was relocated to the end for a cleaner structure.


```{r}
extract_kml_field <- function(html_text, field_name) {
  if (is.na(html_text) || html_text == "") return(NA_character_)
  
  page <- read_html(html_text)
  rows <- page %>% html_elements("tr")
  
  value <- rows %>%
    keep(~ html_text2(html_element(.x, "th")) == field_name) %>%
    html_element("td") %>%
    html_text2()
  
  if (length(value) == 0) NA_character_ else value
}

mpsz <- mpsz %>%
  mutate(
    REGION_N = map_chr(Description, extract_kml_field, "REGION_N"),
    PLN_AREA_N = map_chr(Description, extract_kml_field, "PLN_AREA_N"),
    SUBZONE_N = map_chr(Description, extract_kml_field, "SUBZONE_N"),
    SUBZONE_C = map_chr(Description, extract_kml_field, "SUBZONE_C")
  ) %>%
  select(-Name, -Description) %>%
  relocate(geometry, .after = last_col())
```

```{r}
glimpse(mpsz)
```
## 4.9 Splitting Singapore Map by Regions

This step filters the Singapore planning subzone dataset (mpsz) into five separate regions: Central, East, West, North, and North-East. Each subset can then be used for regional-level spatial analysis and visualization.
```{r}
cr <- mpsz %>%
  filter(REGION_N == "CENTRAL REGION")
er <- mpsz %>%
  filter(REGION_N == "EAST REGION")
wr <- mpsz %>%
  filter(REGION_N == "WEST REGION")
nr <- mpsz %>%
  filter(REGION_N == "NORTH REGION")
ner <- mpsz %>%
  filter(REGION_N == "NORTH-EAST REGION")
```



## Visualising the distribution

```{r}
# Clinics
ggplot(biz_clinic, aes(x = MONTH_ABBR)) +
  geom_bar() +
  facet_wrap(~ YEAR, ncol = 2) +
  labs(title = "Monthly Count of New Clinics, 2024–2025",
       x = "Month", y = "Count")

# Massage
ggplot(biz_Massage, aes(x = MONTH_ABBR)) +
  geom_bar() +
  facet_wrap(~ YEAR, ncol = 2) +
  labs(title = "Monthly Count of New Massage, 2024–2025",
       x = "Month", y = "Count")

# Fitness centres
ggplot(biz_fitness, aes(x = MONTH_ABBR)) +
  geom_bar() +
  facet_wrap(~ YEAR, ncol = 2) +
  labs(title = "Monthly Count of New Fitness Centres , 2024–2025",
       x = "Month", y = "Count")
```

## Visualising the business

```{r}
tmap_mode("view")
```

```{r}
# Clinics
tm_shape(biz_clinic_sf) +
  tm_dots(col = "blue", alpha = 0.7, size = 0.06) +
  tm_view(set.view = c(103.851959, 1.290270, 11)) +
  tm_layout(title = "Locations of New Clinics")

```

```{r}
# Massage
tm_shape(biz_Massage_sf) +
  tm_dots(col = "red", alpha = 0.7, size = 0.3) +
  tm_view(set.view = c(103.851959, 1.290270, 11)) +
  tm_layout(title = "Locations of New Massage")
```

```{r}
# Fitness centres
tm_shape(biz_fitness_sf) +
  tm_dots(col = "green", alpha = 0.7, size = 0.3) +
  tm_view(set.view = c(103.851959, 1.290270, 11)) +
  tm_layout(title = "Locations of New Fitness Centres")
```

```{r}
biz_clinic_sf  <- biz_clinic_sf  %>% mutate(type = "Clinic")
biz_Massage_sf <- biz_Massage_sf %>% mutate(type = "Massage")
biz_fitness_sf <- biz_fitness_sf %>% mutate(type = "Fitness")

biz_all_sf <- rbind(biz_clinic_sf, biz_Massage_sf, biz_fitness_sf)

tmap_mode("view")

tm_shape(biz_all_sf) +
  tm_dots(col = "type", palette = c("blue","red","green"),
          alpha = 0.7, size = 0.2, legend.show = TRUE) +
  tm_view(set.view = c(103.851959, 1.290270, 11)) +
  tm_layout(title = "Locations of Clinics, Massage and Fitness Centres",
            legend.outside = TRUE)
```

```{r}
tmap_mode("plot")
```

## overview for SinMap

```{r}
tmap_style("natural")
tm_shape(mpsz) +
  tm_fill("lightgreen", title = "Singapore Boundary") +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Map of Singapore",
            main.title.position = "center",
            main.title.size = 1.0,
            legend.height = 0.35, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha = 0.2)
```





## overview for 3 industry

::: panel-tabset
# Clinics

```{r}
qtm(biz_clinic_sf)
```

# Massage

```{r}
qtm(biz_Massage_sf)
```

# Fitness Centres

```{r}
qtm(biz_fitness_sf)
```
:::

## region overview for 3 industry

::: panel-tabset
# Clinics

```{r}
tm_shape(mpsz) +
  tm_polygons("REGION_N", alpha = 0.3) +
tm_shape(biz_clinic_sf) +
  tm_dots(col = "blue", size = 0.05, popup.vars = c("entity_name", "ADDRESS"))
```

# Massage

```{r}
tm_shape(mpsz) +
  tm_polygons("REGION_N", alpha = 0.3) +
tm_shape(biz_Massage_sf) +
  tm_dots(col = "blue", size = 0.05, popup.vars = c("entity_name", "ADDRESS"))
```

# Fitness Centres

```{r}
tm_shape(mpsz) +
  tm_polygons("REGION_N", alpha = 0.3) +
tm_shape(biz_fitness_sf) +
  tm_dots(col = "blue", size = 0.05, popup.vars = c("entity_name", "ADDRESS"))
```
:::

## choropleth map

```{r}
mpsz <- st_zm(mpsz, drop = TRUE)     
mpsz <- st_make_valid(mpsz)          
mpsz <- st_collection_extract(mpsz, "POLYGON")
mpsz <- st_cast(mpsz, "MULTIPOLYGON", warn = FALSE)
mpsz <- st_transform(mpsz, 3414)     
stopifnot(all(st_is_valid(mpsz)))
```

::: panel-tabset
# Clinics

```{r}
clinic_with_zone <- st_join(biz_clinic_sf, mpsz, join = st_within)

clinic_count <- clinic_with_zone %>%
  st_drop_geometry() %>%
  group_by(SUBZONE_N) %>%
  summarise(count = n())

mpsz_clinic <- mpsz %>%
  left_join(clinic_count, by = "SUBZONE_N")

tm_shape(mpsz_clinic) +
  tm_polygons("count", palette = "Blues", title = "Clinics") +
tm_shape(biz_clinic_sf) +
  tm_dots(col = "blue", size = 0.05)
```

# Massage

```{r}
massage_with_zone <- st_join(biz_Massage_sf, mpsz, join = st_within)

massage_count <- massage_with_zone %>%
  st_drop_geometry() %>%
  group_by(SUBZONE_N) %>%
  summarise(count = n())

mpsz_massage <- mpsz %>%
  left_join(massage_count, by = "SUBZONE_N")

tm_shape(mpsz_massage) +
  tm_polygons("count", palette = "Blues", title = "Massage") +
tm_shape(biz_Massage_sf) +
  tm_dots(col = "red", size = 0.05)
```

# Fitness Centres

```{r}
fitness_with_zone <- st_join(biz_fitness_sf, mpsz, join = st_within)

fitness_count <- fitness_with_zone %>%
  st_drop_geometry() %>%
  group_by(SUBZONE_N) %>%
  summarise(count = n())

mpsz_fitness <- mpsz %>%
  left_join(fitness_count, by = "SUBZONE_N")

tm_shape(mpsz_fitness) +
  tm_polygons("count", palette = "Blues", title = "Fitness Businesses") +
tm_shape(biz_fitness_sf) +
  tm_dots(col = "red", size = 0.05)
```
:::

## Creating owin object

```{r}
sg_owin <- as.owin(mpsz)
```

```{r}
class(sg_owin)
```

```{r}
plot(sg_owin)
```

## Converting sf data frames to ppp class

::: panel-tabset
# Clinics

```{r}
clinic_ppp <- as.ppp(biz_clinic_sf)
class(clinic_ppp)
```

# Massage

```{r}
Massage_ppp <- as.ppp(biz_Massage_sf)
class(Massage_ppp)
```

# Fitness Centres

```{r}
fitness_ppp <- as.ppp(biz_fitness_sf)
class(fitness_ppp)
```
:::

## Combining point events object and owin object

::: panel-tabset
# Clinics

```{r}
clinic_ppp = clinic_ppp[sg_owin]
```

```{r}
clinic_ppp
```

# Massage

```{r}
Massage_ppp = Massage_ppp[sg_owin]
```

```{r}
Massage_ppp
```

# Fitness Centres

```{r}
fit_ppp = fitness_ppp[sg_owin]
```

```{r}
fit_ppp
```
:::

## Clark-Evan Test for Nearest Neighbour Analysis

Nearest Neighbor Analysis (NNA) is a spatial statistics method that calculates the average distance between each point and its closest neighbor to determine if a pattern of points is clustered, dispersed, or randomly distributed.

Clark-Evans test is a specific statistical method used within NNA to quantify whether a point pattern is clustered, random, or uniformly spaced, using the Clark-Evans aggregation index (R) to describe this pattern. NNA provides a numerical value that describes the degree of clustering or regularity, and the Clark-Evans test calculates a specific index (R) for this purpose.

::: panel-tabset
# Clinics

假设检验

```{r}
clarkevans.test(clinic_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"))
```

# Massage

```{r}
clarkevans.test(Massage_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"))
```

# Fitness Centres

```{r}
clarkevans.test(fit_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"))
```
:::

## Kernel Density Estimation Method

Kernel Density Estimation (KDE) is a valuable tool for visualising and analyzing first-order spatial point patterns. It is widely considered a method within Exploratory Spatial Data Analysis (ESDA) because it’s used to visualize and understand spatial data patterns by transforms discrete point data (like locations of childcare service, crime incidents or disease cases) into continuous density surfaces that reveal clusters and variations in event occurrences, without making prior assumptions about data distribution. It helps to begin understanding data distribution, identify hotspots, and explore relationships between spatial variables before performing more rigorous analysis.

### automatic bandwidth selection method

::: panel-tabset
# Clinics

```{r}
clinic_ppp_km <- rescale.ppp(
  clinic_ppp, 1000, "km")

kde_clinic_diggle_km <- density(clinic_ppp_km,
                                sigma = bw.diggle,
                                edge = TRUE,
                                kernel = "gaussian")

plot(kde_clinic_diggle_km, main = "Clinic KDE (bw.diggle, km)")
```

# Massage

```{r}
massage_ppp_km <- rescale.ppp(
  Massage_ppp, 1000, "km")

kde_massage_diggle_km <- density(massage_ppp_km,
                                 sigma = bw.diggle,
                                 edge = TRUE,
                                 kernel = "gaussian")

plot(kde_massage_diggle_km, main = "Massage KDE (bw.diggle, km)")
```

# Fitness Centres

```{r}
fit_ppp_km <- rescale.ppp(
  fit_ppp, 1000, "km")

kde_fit_diggle_km <- density(fit_ppp_km,
                              sigma=bw.diggle,
                              edge=TRUE,
                              kernel="gaussian")
plot(kde_fit_diggle_km)
```
:::

### adaptive bandwidth

::: panel-tabset
# Clinics

```{r}
kde_clinic_ab <- adaptive.density(
  clinic_ppp_km, 
  method="kernel")
plot(kde_clinic_ab)
```

# Massage

```{r}
kde_massage_ab <- adaptive.density(
  massage_ppp_km, 
  method="kernel")
plot(kde_massage_ab)
```

# Fitness Centres

```{r}
kde_fit_ab <- adaptive.density(
  fit_ppp_km, 
  method="kernel")
plot(kde_fit_ab)
```
:::

### cartographic quality KDE map


```{r}
kde_clinic_bw_terra <- rast(kde_clinic_diggle_km)
crs(kde_clinic_bw_terra) <- "EPSG:3414"
class(kde_clinic_bw_terra)
kde_clinic_bw_terra

kde_massage_bw_terra <- rast(kde_massage_diggle_km)
crs(kde_massage_bw_terra) <- "EPSG:3414"
class(kde_massage_bw_terra)
kde_massage_bw_terra

kde_fit_bw_terra <- rast(kde_fit_diggle_km)
crs(kde_fit_bw_terra) <- "EPSG:3414"
class(kde_fit_bw_terra)
kde_fit_bw_terra
```

::: panel-tabset
# Clinics

```{r}
tm_shape(kde_clinic_bw_terra) + 
  tm_raster(col.scale = 
              tm_scale_continuous(
                values = "viridis"),
            col.legend = tm_legend(
            title = "Density values",
            title.size = 0.7,
            text.size = 0.7,
            bg.color = "white",
            bg.alpha = 0.7,
            position = tm_pos_in(
              "right", "bottom"),
            frame = TRUE)) +
  tm_graticules(labels.size = 0.7) +
  tm_compass() +
  tm_layout(scale = 1.0)
```

# Massage

```{r}
tm_shape(kde_massage_bw_terra) + 
  tm_raster(col.scale = 
              tm_scale_continuous(
                values = "viridis"),
            col.legend = tm_legend(
            title = "Density values",
            title.size = 0.7,
            text.size = 0.7,
            bg.color = "white",
            bg.alpha = 0.7,
            position = tm_pos_in(
              "right", "bottom"),
            frame = TRUE)) +
  tm_graticules(labels.size = 0.7) +
  tm_compass() +
  tm_layout(scale = 1.0)
```

# Fitness Centres

```{r}
tm_shape(kde_fit_bw_terra) + 
  tm_raster(col.scale = 
              tm_scale_continuous(
                values = "viridis"),
            col.legend = tm_legend(
            title = "Density values",
            title.size = 0.7,
            text.size = 0.7,
            bg.color = "white",
            bg.alpha = 0.7,
            position = tm_pos_in(
              "right", "bottom"),
            frame = TRUE)) +
  tm_graticules(labels.size = 0.7) +
  tm_compass() +
  tm_layout(scale = 1.0)
```
:::

# region



```{r}
par(mfrow=c(2,2))
plot(st_geometry(cr), main = "CENTRAL REGION")
plot(st_geometry(er), main = "EAST REGION")
plot(st_geometry(wr), main = "WEST REGION")
plot(st_geometry(nr), main = "JNORTH REGION")
plot(st_geometry(ner), main = "NORTH-EAST REGION")
```

```{r}
cr_owin = as.owin(cr)
er_owin = as.owin(er)
wr_owin = as.owin(wr)
nr_owin = as.owin(nr)
ner_owin = as.owin(ner)
```

## Group by Region

::: panel-tabset
# Clinics

```{r}
clinic_cr_ppp  = clinic_ppp[cr_owin]
clinic_er_ppp  = clinic_ppp[er_owin]
clinic_wr_ppp  = clinic_ppp[wr_owin]
clinic_nr_ppp  = clinic_ppp[nr_owin]
clinic_ner_ppp = clinic_ppp[ner_owin]

clinic_cr_ppp.km  = rescale.ppp(clinic_cr_ppp,  1000, "km")
clinic_er_ppp.km  = rescale.ppp(clinic_er_ppp,  1000, "km")
clinic_wr_ppp.km  = rescale.ppp(clinic_wr_ppp,  1000, "km")
clinic_nr_ppp.km  = rescale.ppp(clinic_nr_ppp,  1000, "km")
clinic_ner_ppp.km = rescale.ppp(clinic_ner_ppp, 1000, "km")

par(mfrow = c(2, 3), cex = 0.3)
plot(unmark(clinic_cr_ppp.km),  main = "CENTRAL REGION (Clinic)")
plot(unmark(clinic_er_ppp.km),  main = "EAST REGION (Clinic)")
plot(unmark(clinic_wr_ppp.km),  main = "WEST REGION (Clinic)")
plot(unmark(clinic_nr_ppp.km),  main = "NORTH REGION (Clinic)")
plot(unmark(clinic_ner_ppp.km), main = "NORTH-EAST REGION (Clinic)")
```

# Massage

```{r}
massage_cr_ppp  = Massage_ppp[cr_owin]
massage_er_ppp  = Massage_ppp[er_owin]
massage_wr_ppp  = Massage_ppp[wr_owin]
massage_nr_ppp  = Massage_ppp[nr_owin]
massage_ner_ppp = Massage_ppp[ner_owin]

massage_cr_ppp.km  = rescale.ppp(massage_cr_ppp,  1000, "km")
massage_er_ppp.km  = rescale.ppp(massage_er_ppp,  1000, "km")
massage_wr_ppp.km  = rescale.ppp(massage_wr_ppp,  1000, "km")
massage_nr_ppp.km  = rescale.ppp(massage_nr_ppp,  1000, "km")
massage_ner_ppp.km = rescale.ppp(massage_ner_ppp, 1000, "km")

par(mfrow = c(2, 3), cex = 0.3)
plot(unmark(massage_cr_ppp.km),  main = "CENTRAL REGION (Massage)")
plot(unmark(massage_er_ppp.km),  main = "EAST REGION (Massage)")
plot(unmark(massage_wr_ppp.km),  main = "WEST REGION (Massage)")
plot(unmark(massage_nr_ppp.km),  main = "NORTH REGION (Massage)")
plot(unmark(massage_ner_ppp.km), main = "NORTH-EAST REGION (Massage)")
```

# Fitness Centres

```{r}
fit_cr_ppp = fitness_ppp[cr_owin]
fit_er_ppp = fitness_ppp[er_owin]
fit_wr_ppp = fitness_ppp[wr_owin]
fit_nr_ppp = fitness_ppp[nr_owin]
fit_ner_ppp = fitness_ppp[ner_owin]

fit_cr_ppp.km = rescale.ppp(fit_cr_ppp, 1000, "km")
fit_er_ppp.km = rescale.ppp(fit_er_ppp, 1000, "km")
fit_wr_ppp.km = rescale.ppp(fit_wr_ppp, 1000, "km")
fit_nr_ppp.km = rescale.ppp(fit_nr_ppp, 1000, "km")
fit_ner_ppp.km = rescale.ppp(fit_ner_ppp, 1000, "km")

par(mfrow=c(2,2), cex = 0.3)
plot(unmark(fit_cr_ppp.km), 
  main="CENTRAL REGIONl")
plot(unmark(fit_er_ppp.km), 
  main="EAST REGION")
plot(unmark(fit_wr_ppp.km), 
  main="WEST REGION")
plot(unmark(fit_nr_ppp.km), 
  main="JNORTH REGION")
plot(unmark(fit_ner_ppp.km), 
  main="NORTH-EAST REGION")
```
:::

# Second-order Spatial Point Patterns Analysis

```{r}
set.seed(1234)
```

## clinic

假设检验

::: panel-tabset
# G-function

```{r}
G_clinic_sg = Gest(clinic_ppp, correction = "border")
plot(G_clinic_sg, xlim=c(0,500))
```

```{r}
#| cache: true
G_clinic_sg.csr <- envelope(clinic_ppp, Gest, nsim = 999)
```

```{r}
plot(G_clinic_sg.csr)
```

# F-function

```{r}
F_clinic_sg = Fest(clinic_ppp)
plot(F_clinic_sg)
```

```{r}
#| cache: true
F_clinic_sg.csr <- envelope(clinic_ppp, Fest, nsim = 999)
```

```{r}
plot(F_clinic_sg.csr)
```

# K-function

```{r}
K_clinic_sg = Kest(clinic_ppp, correction = "Ripley")
plot(K_clinic_sg, . -r ~ r, ylab= "K(d)-r", xlab = "d(m)")
```

```{r}
#| cache: true
K_clinic_sg.csr <- envelope(clinic_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(K_clinic_sg.csr, . - r ~ r, xlab="d", ylab="K(d)-r")
```

# L-function

```{r}
L_clinic_sg = Lest(clinic_ppp, correction = "Ripley")
plot(L_clinic_sg, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)")
```

```{r}
#| cache: true
L_clinic_sg.csr <- envelope(clinic_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(L_clinic_sg.csr, . - r ~ r, xlab="d", ylab="L(d)-r")
```
:::

## massage

假设检验

::: panel-tabset
# G-function

```{r}
G_Massage_sg = Gest(Massage_ppp, correction = "border")
plot(G_Massage_sg, xlim=c(0,500))
```

```{r}
#| cache: true
G_Massage_sg.csr <- envelope(Massage_ppp, Gest, nsim = 999)
```

```{r}
plot(G_Massage_sg.csr)
```

# F-function

```{r}
F_Massage_sg = Fest(Massage_ppp)
plot(F_Massage_sg)
```

```{r}
#| cache: true
F_Massage_sg.csr <- envelope(Massage_ppp, Fest, nsim = 999)
```

```{r}
plot(F_Massage_sg.csr)
```

# K-function

```{r}
K_Massage_sg = Kest(Massage_ppp, correction = "Ripley")
plot(K_Massage_sg, . -r ~ r, ylab= "K(d)-r", xlab = "d(m)")
```

```{r}
#| cache: true
K_Massage_sg.csr <- envelope(Massage_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(K_Massage_sg.csr, . - r ~ r, xlab="d", ylab="K(d)-r")
```

# L-function

```{r}
L_Massage_sg = Lest(Massage_ppp, correction = "Ripley")
plot(L_Massage_sg, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)")
```

```{r}
#| cache: true
L_Massage_sg.csr <- envelope(Massage_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(L_Massage_sg.csr, . - r ~ r, xlab="d", ylab="L(d)-r")
```
:::

## fitness

假设检验

::: panel-tabset
# G-function

```{r}
G_fit_sg = Gest(fit_ppp, correction = "border")
plot(G_fit_sg, xlim=c(0,500))
```

```{r}
#| cache: true
G_fit_sg.csr <- envelope(fit_ppp, Gest, nsim = 999)
```

```{r}
plot(G_fit_sg.csr)
```

# F-function

```{r}
F_fit_sg = Fest(fit_ppp)
plot(F_fit_sg)
```

```{r}
#| cache: true
F_fit_sg.csr <- envelope(fit_ppp, Fest, nsim = 999)
```

```{r}
plot(F_fit_sg.csr)
```

# K-function

```{r}
K_fit_sg = Kest(fit_ppp, correction = "Ripley")
plot(K_fit_sg, . -r ~ r, ylab= "K(d)-r", xlab = "d(m)")
```

```{r}
#| cache: true
K_fit_sg.csr <- envelope(fit_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(K_fit_sg.csr, . - r ~ r, xlab="d", ylab="K(d)-r")
```

# L-function

```{r}
L_fit_sg = Lest(fit_ppp, correction = "Ripley")
plot(L_fit_sg, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)")
```

```{r}
#| cache: true
L_fit_sg.csr <- envelope(fit_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(L_fit_sg.csr, . - r ~ r, xlab="d", ylab="L(d)-r")
```
:::
