---
title: "Take-home Exercise 1"
author: "Li Yuquan"
description: ""
date: "September 15, 2025"
date-modified: "last-modified"
format: html
editor: visual
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# 1 Overview

## 1.1 Objectives

The location choices of businesses play a pivotal role in shaping urban development, influencing economic growth, job creation, and the provision of resources and services. Strategic siting not only creates competitive advantages and supports efficient supply chains, but also contributes to a city’s identity through specialized districts and accessible infrastructure.

Studying the spatio-temporal patterns of new businesses is therefore essential, as it highlights how economic activity and urban forms evolve over time. Such analysis enables urban planners to anticipate growth, identify clusters and trends, improve resource allocation, and design more sustainable and livable cities.

In this exercise, the focus is on new business entities established in Singapore during the first six months of 2025. The objectives are:

• To analyze the “where” and “when” of new business establishments, uncovering spatial clusters, temporal trends, and development hotspots.

• To explore complex interaction patterns beyond simple density measures, such as the clustering of related firms, path-dependency of business locations, and structural changes in commercial areas.

## 1.2 The Task

1.  Download ACRA (Accounting and Corporate Regulatory Authority) Information on Corporate Entities datasets from data.gov.sg.

2.  Select at least three business types by referring to [2-digit or 3-digit Singapore Standard Industrial Classification (SSIS) 2020](https://classification.codes/classifications/industry/ssic-singapore). Prepare and tidy the data by applying appropriate data wrangling method.

3.  Extract the selected business entities registered between 1st January 2024 to 30th June 2025.

4.  Geocode the selected records and convert them into simple point features.

5.  Compute spatial and spatio-temporal KDE of the selected business types.

6.  Perform second order spatial and spatio-temporal analysis on the selected business types.

7.  Describe and discuss the analysis results obtained from Step 5 and 6.

8.  For each geovisualisation prepared, write a short report of not more than 150 words describing the spatio-temporal patterns revealed by the geovisualisation and statistical graphics.

9.  With reference to the analysis results above, discuss how these findings can be used to support urban land use planning and management. (Not more than 500 words)

# 1.3 R packages

| **Package** | **Description** |
|------------------------------------|------------------------------------|
| sf | Simple Features for handling and analyzing spatial vector data. |
| spNetwork | Network-based spatial analysis, e.g. kernel density along road networks. |
| spatstat | Spatial point pattern analysis (ppp, owin, G/F/K/L functions, KDE). |
| tmap | Thematic maps creation (static and interactive). |
| tidyverse | Data science suite (dplyr, ggplot2, readr, purrr, etc.). |
| httr | Tools for working with HTTP to access web APIs. |
| rvest | Web scraping and HTML parsing. |
| terra | Spatial raster data handling and analysis. |
| lwgeom | Geometry operations, fixing invalid shapes (wraps GEOS). |

```{r}
pacman::p_load(sf, spNetwork, spatstat,tmap, tidyverse, httr, rvest, terra, lwgeom, stpp, dplyr, lubridate, spatstat.geom, sf, spatstat, sparr, 
               tmap, stpp, tidyverse, magick, fields)
```

# 1.4 The Data

| **Dataset Name** | **Description** | **Format** | **Source** |
|------------------|------------------|------------------|------------------|
| ACRA Information on Corporate Entities ('A-Z-Others') | Business registration information (clinics, massage, fitness filtered by SSIC) | CSV | data.gov.sg |
| OneMap Geocoding Results | Geocoded coordinates (X, Y) for postal codes | API | [OneMap API (Singapore Land Authority)](https://onemap.gov.sg/api/common/elastic/search) |
| MasterPlan2019SubzoneBoundaryNoSeaKML | Planning region and subzone polygon boundaries | KML | URA (Urban Redevelopment Authority) |
| Singapore Standard Industrial Classification (SSIS) 2020 | SSIS Code Matching Table | Excel | [2-digit or 3-digit Singapore Standard Industrial Classification (SSIS) 2020](https://classification.codes/classifications/industry/ssic-singapore) |

# 4 Data Import and Wrangling

## 4.1 Importing ACRA data

Import ACRA data to R studio, combin all 27 file togeter and save it as rds file.

```{r}

folder_path <- "data/aspatial"
file_list <- list.files(path = folder_path,
                        pattern = "^ACRA*.*\\.csv$",
                        full.names = TRUE)

acra_data <- file_list %>%
  map_dfr(read_csv)
```

```{r}
write_rds(acra_data,
          "data/rds/acra_data.rds")

```

```{r}
summary(acra_data) 
```

```{r}
glimpse(acra_data)
```

## 4.2 Importing MasterPlan2019SubzoneBoundaryNoSea data

The Master Plan 2019 Subzone Boundary (KML) file was imported, re-projected to Singapore’s SVY21 coordinate system (EPSG:3414), and saved as an RDS file for later mapping and analysis.

```{r}

mpsz <- st_read("data/MasterPlan2019SubzoneBoundaryNoSeaKML.kml")
mpsz <- mpsz %>% st_transform(crs=3414)

```

```{r}
write_rds(mpsz, "data/rds/mpsz.rds")
mpsz <- read_rds("data/rds/mpsz.rds")
```

```{r}
summary(mpsz) 
```

```{r}
glimpse(mpsz)
```

## 4.3 Tidying ACRA data

The ACRA dataset was filtered to keep only four SSIC codes representing three industry types: Clinics (86201, 86202), Massage services (96024), and Fitness centres (93111). The data was cleaned (date fields, year/month extraction, postal code formatting) and then split into three subsets: biz_clinic, biz_Massage, and biz_fitness.

| **Industry Type** | **SSIC Code(s)** | **Description** |
|------------------------|------------------------|------------------------|
| Clinics | 86201, 86202 | Clinics and other general medical services (Western & Non-Western) |
| Massage Services | 96024 | Massage parlours and foot reflexology |
| Fitness Centres | 93111 | Fitness centres and gymnasiums |

```{r}
ssic_vec <- c("86201", "86202", "96024", "93111")

acra_core <- acra_data %>%
  select(1:24) %>% 
  mutate(
    primary_ssic_code = as.character(primary_ssic_code),
    date  = as.Date(registration_incorporation_date),
    YEAR  = year(date),
    MONTH_NUM  = month(date),
    MONTH_ABBR = month(date, label = TRUE, abbr = TRUE)
  ) %>%
  filter(
    date >= as.Date("2024-01-01"),
    date <= as.Date("2025-06-30"),
    primary_ssic_code %in% ssic_vec
  ) %>%
  mutate(
    postal_code = str_pad(postal_code, width = 6, side = "left", pad = "0")
  )


biz_clinic  <- acra_core %>% filter(primary_ssic_code %in% c("86201", "86202"))
biz_Massage  <- acra_core %>% filter(primary_ssic_code %in% c("96024"))
biz_fitness <- acra_core %>% filter(primary_ssic_code == "93111")
```

## 4.4 Geocoding

The postal codes from the three business subsets (clinics, massage, fitness centres) were extracted and cleaned to remove duplicates and missing values. These unique postal codes were then sent to the OneMap API for geocoding. For each code, the API returned geographic coordinates (X, Y) and address details if found. The results were stored in a found dataset, while unmatched postal codes were recorded separately in a not_found dataset.

```{r}
postcodes <- c(biz_clinic$postal_code,
               biz_Massage$postal_code,
               biz_fitness$postal_code) %>%
  unique() %>%
  na.omit()

url <- "https://onemap.gov.sg/api/common/elastic/search"

found     <- data.frame()
not_found <- data.frame(postcode = character())

for (pc in postcodes) {
  query <- list(
    searchVal = pc,
    returnGeom = "Y",
    getAddrDetails = "Y",
    pageNum = "1"
  )
  res  <- GET(url, query = query)
  json <- content(res)
  
  if (!is.null(json$found) && json$found != 0) {
    df <- as.data.frame(json$results, stringsAsFactors = FALSE)
    df$input_postcode <- pc
    found <- bind_rows(found, df)
  } else {
    not_found <- bind_rows(not_found, data.frame(postcode = pc))
  }
 
}
```

## 4.5 Tidying the geocoded data

The geocoding results were cleaned by selecting only the first 10 relevant fields from the found dataset to keep essential location attributes.

```{r}
found <- found %>%
  select(1:10)
```

## 4.6 Appending the location information

The cleaned geocoded results were joined back to the three business datasets (biz_clinic, biz_Massage, biz_fitness) using postal code as the key, enriching each business record with its spatial coordinates and address details. The updated datasets were then saved as RDS files for later use.

```{r}

biz_clinic  <- biz_clinic  %>% left_join(found, by = c("postal_code" = "POSTAL"))
biz_Massage  <- biz_Massage  %>% left_join(found, by = c("postal_code" = "POSTAL"))
biz_fitness <- biz_fitness %>% left_join(found, by = c("postal_code" = "POSTAL"))
```

```{r}
write_rds(biz_clinic, "data/rds/biz_clinic.rds")
write_rds(biz_Massage, "data/rds/biz_Massage.rds")
write_rds(biz_fitness, "data/rds/biz_fitness.rds")


```

## 4.7 Converting into SF data frame

Each business dataset was transformed into an sf (simple features) object (biz_clinic_sf, biz_Massage_sf, biz_fitness_sf), keeping only records with valid coordinates (X, Y). The coordinate system was set to EPSG:3414 (SVY21 / Singapore TM), enabling spatial mapping and analysis.

```{r}
biz_clinic_sf <- st_as_sf(
  biz_clinic %>% filter(!is.na(X), !is.na(Y)),
  coords = c("X","Y"), crs = 3414
)

biz_Massage_sf <- st_as_sf(
  biz_Massage %>% filter(!is.na(X), !is.na(Y)),
  coords = c("X","Y"), crs = 3414
)

biz_fitness_sf <- st_as_sf(
  biz_fitness %>% filter(!is.na(X), !is.na(Y)),
  coords = c("X","Y"), crs = 3414
)
```

### 4.7.1 Glimpse for 3 industry

By performing the Glimpse operation on the data sets of the three industries, we can determine the amount of data contained in each one by looking at the number of rows.

::: panel-tabset
# Clinics

```{r}
glimpse(biz_clinic_sf)
```

# Massage

```{r}
glimpse(biz_Massage_sf)
```

# Fitness Centres

```{r}
glimpse(biz_fitness_sf)
```
:::

## 4.8 Extracting Attributes from KML Description

The imported Master Plan 2019 Subzone Boundary (mpsz) dataset contained key attributes (e.g., region, planning area, subzone name, subzone code) embedded inside the Description field of the KML file. A custom function extract_kml_field() was created to parse the HTML structure and extract these values. The extracted fields (REGION_N, PLN_AREA_N, SUBZONE_N, SUBZONE_C) were then added as new columns to the dataset. Finally, redundant fields (Name, Description) were removed, and the geometry column was relocated to the end for a cleaner structure.

```{r}
extract_kml_field <- function(html_text, field_name) {
  if (is.na(html_text) || html_text == "") return(NA_character_)
  
  page <- read_html(html_text)
  rows <- page %>% html_elements("tr")
  
  value <- rows %>%
    keep(~ html_text2(html_element(.x, "th")) == field_name) %>%
    html_element("td") %>%
    html_text2()
  
  if (length(value) == 0) NA_character_ else value
}

mpsz <- mpsz %>%
  mutate(
    REGION_N = map_chr(Description, extract_kml_field, "REGION_N"),
    PLN_AREA_N = map_chr(Description, extract_kml_field, "PLN_AREA_N"),
    SUBZONE_N = map_chr(Description, extract_kml_field, "SUBZONE_N"),
    SUBZONE_C = map_chr(Description, extract_kml_field, "SUBZONE_C")
  ) %>%
  select(-Name, -Description) %>%
  relocate(geometry, .after = last_col())
```

```{r}
glimpse(mpsz)
```

# 5 Exploratory Visualization

In this section, I will conduct a preliminary visualization and exploratory analysis on the processed data.

## 5.1 Visualising the distribution

::: panel-tabset
# Clinics

The number of clinic registrations in 2024 generally remained between 20 and 35 per month, with a noticeable peak in the middle of the year (May–July), reaching or exceeding 35. In the first half of 2025, the trend remained strong, with registrations in May climbing above 40, the highest value during the observation period. This suggests that the clinic industry exhibits clear seasonal concentration in registrations, with stronger growth momentum in 2025.

```{r}
# Clinics
ggplot(biz_clinic, aes(x = MONTH_ABBR)) +
  geom_bar() +
  facet_wrap(~ YEAR, ncol = 2) +
  labs(title = "Monthly Count of New Clinics, 2024–2025",
       x = "Month", y = "Count")

```

# Massage

The number of massage-related business registrations in 2024 was relatively balanced across months, mostly between 20 and 30, with little fluctuation. In the first half of 2025, registrations continued at a similar level, with small peaks above 30 in January and March 2024 and in May 2025. Overall, this industry demonstrates a stable and evenly distributed pattern without obvious seasonal fluctuations.

```{r}
# Massage
ggplot(biz_Massage, aes(x = MONTH_ABBR)) +
  geom_bar() +
  facet_wrap(~ YEAR, ncol = 2) +
  labs(title = "Monthly Count of New Massage, 2024–2025",
       x = "Month", y = "Count")

```

# Fitness Centres

The number of fitness centre registrations fluctuated widely during 2024–2025. In May 2025, a clear peak was observed with 40 new businesses, while in most other months the number hovered between 15 and 25, showing significant variation. This indicates that the fitness industry experiences instability in the number of new company registrations.

```{r}
# Fitness centres
ggplot(biz_fitness, aes(x = MONTH_ABBR)) +
  geom_bar() +
  facet_wrap(~ YEAR, ncol = 2) +
  labs(title = "Monthly Count of New Fitness Centres , 2024–2025",
       x = "Month", y = "Count")
```
:::

## 5.2 Spatial Distribution of Businesses

```{r}
tmap_mode("view")
```

::: panel-tabset
# Clinics

The distribution of clinics is relatively balanced, covering the entire island, from the city centre to outer residential districts, though overall slightly more in the east than in the west. Compared to massage and fitness centres, clinics are closer to a public service function, with wider coverage to ensure residents in different communities have access to basic healthcare. As a result, evenly spread blue dots can also be observed in western, northern, and eastern neighbourhoods, highlighting the ubiquity and accessibility of clinics.

```{r}
tm_shape(biz_clinic_sf) +
  tm_dots(col = "blue", alpha = 0.7, size = 0.06) +
  tm_view(set.view = c(103.851959, 1.290270, 11)) +
  tm_layout(title = "Locations of New Clinics")
```

# Massage

Massage businesses are more concentrated within the city area, especially in Orchard Road, Bugis, and the central-southern regions, where red dots are densely packed. This pattern is consistent with the industry’s reliance on areas with high foot traffic and active consumer spending. In the outer areas, the points are relatively fewer, showing that massage services mainly target the city centre and certain prosperous residential zones, with a strong dependence on commercial demand.

```{r}
tm_shape(biz_Massage_sf) +
  tm_dots(col = "red", alpha = 0.7, size = 0.3) +
  tm_view(set.view = c(103.851959, 1.290270, 11)) +
  tm_layout(title = "Locations of New Massage")
```

# Fitness Centres

The spatial distribution of fitness centres is relatively extensive, stretching from the city centre to residential areas in both the east and west of Singapore, with a higher concentration in the eastern and central parts. This suggests that the fitness industry not only serves the central business districts but also penetrates into neighbourhoods, providing sports and leisure facilities for the community. The clusters around the Harbourfront area and new residential developments in the east also reflect the alignment of fitness consumption with emerging housing areas.

```{r}
tm_shape(biz_fitness_sf) +
  tm_dots(col = "green", alpha = 0.7, size = 0.3) +
  tm_view(set.view = c(103.851959, 1.290270, 11)) +
  tm_layout(title = "Locations of New Fitness Centres")
```

# Combine

This map shows the overall distribution of the three industries across Singapore. The points are mainly concentrated along major residential and commercial areas. In particular, the central and southern regions (such as the city centre and Orchard area) are densely clustered, indicating that clinics, massage parlours, and fitness centres tend to locate in core zones to serve areas with high population density and strong consumption power. Peripheral regions also have some distribution, but the points are noticeably sparser.

```{r}
biz_clinic_sf  <- biz_clinic_sf  %>% mutate(type = "Clinic")
biz_Massage_sf <- biz_Massage_sf %>% mutate(type = "Massage")
biz_fitness_sf <- biz_fitness_sf %>% mutate(type = "Fitness")

biz_all_sf <- rbind(biz_clinic_sf, biz_Massage_sf, biz_fitness_sf)

tmap_mode("view")

tm_shape(biz_all_sf) +
  tm_dots(col = "type", palette = c("blue","red","green"),
          alpha = 0.7, size = 0.2, legend.show = TRUE) +
  tm_view(set.view = c(103.851959, 1.290270, 11)) +
  tm_layout(title = "Locations of Clinics, Massage and Fitness Centres",
            legend.outside = TRUE)
```
:::

```{r}
tmap_mode("plot")
```

## 5.3 Singapore Basemap Overview

The map above presents the administrative subzone boundaries of Singapore as defined in the Urban Redevelopment Authority (URA) Master Plan 2019. Each polygon represents a planning subzone, which together form the basic spatial units for urban and statistical analysis. The map is projected in the SVY21 / Singapore TM coordinate system (EPSG:3414), which is the national standard for spatial data in Singapore.

This base map provides the essential geographic framework for subsequent spatial analyses. It allows for the overlay of business locations, demographic data, and other point or area-based information, enabling interpretation of spatial patterns across Singapore’s Central, East, West, North, and North-East regions. The inclusion of a scale bar and compass rose ensures that spatial orientation and measurement can be maintained in the analysis.

```{r}
tmap_style("natural")
tm_shape(mpsz) +
  tm_fill("lightgreen", title = "Singapore Boundary") +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Map of Singapore",
            main.title.position = "center",
            main.title.size = 1.0,
            legend.height = 0.35, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha = 0.2)
```

## 5.4 Choropleth Maps by Subzone

```{r}
mpsz <- st_zm(mpsz, drop = TRUE)     
mpsz <- st_make_valid(mpsz)          
mpsz <- st_collection_extract(mpsz, "POLYGON")
mpsz <- st_cast(mpsz, "MULTIPOLYGON", warn = FALSE)
mpsz <- st_transform(mpsz, 3414)     
stopifnot(all(st_is_valid(mpsz)))
```

```{r}
tmap_mode("view")
```

-   Fitness centres are moderately dispersed, balancing between city and neighbourhood coverage.

-   Massage services are highly centralised, clustered in commercial and entertainment hubs, with limited suburban presence.

-   Clinics are the most evenly spread and most numerous, with strong representation across nearly all regions, reflecting their role as a public necessity rather than a lifestyle or leisure option.

In conclusion, based on the content in the chart, two concentrated areas were selected for in-depth analysis for each of the three industry types. For clinics, NOVENA and GAYLANG were chosen. For massage, OUTRAM and KALLANG were selected. For fitness, DOWNTOWN CORE and GAYLANG were chosen.

::: panel-tabset
# Clinics

Clinics exhibit the widest spatial coverage, appearing consistently across central, eastern, and suburban regions. Certain subzones, especially in the eastern part of the island, record over 20 or even 30 clinics, far exceeding the densities observed for fitness or massage services. This balanced yet intensive distribution pattern aligns with the essential service role of clinics, ensuring healthcare accessibility across both densely populated and peripheral residential areas.

```{r}
clinic_with_zone <- st_join(biz_clinic_sf, mpsz, join = st_within)

clinic_count <- clinic_with_zone %>%
  st_drop_geometry() %>%
  group_by(SUBZONE_N) %>%
  summarise(count = n())

mpsz_clinic <- mpsz %>%
  left_join(clinic_count, by = "SUBZONE_N")

tm_shape(mpsz_clinic) +
  tm_polygons("count", palette = "Blues", title = "Clinics") +
tm_shape(biz_clinic_sf) +
  tm_dots(col = "blue", size = 0.05)
```

# Massage

Massage businesses display a more uneven distribution, with denser clusters in the central and southern subzones (such as the city centre and Orchard area). Several subzones register 10 or more businesses, while many peripheral subzones have very few or none. This reflects the industry’s dependence on high-traffic commercial and leisure districts, highlighting its strong commercial orientation compared to fitness businesses.

```{r}
massage_with_zone <- st_join(biz_Massage_sf, mpsz, join = st_within)

massage_count <- massage_with_zone %>%
  st_drop_geometry() %>%
  group_by(SUBZONE_N) %>%
  summarise(count = n())

mpsz_massage <- mpsz %>%
  left_join(massage_count, by = "SUBZONE_N")

tm_shape(mpsz_massage) +
  tm_polygons("count", palette = "Blues", title = "Massage") +
tm_shape(biz_Massage_sf) +
  tm_dots(col = "red", size = 0.05)
```

# Fitness Centres

Fitness businesses are distributed across many subzones but generally show moderate clustering in the central and eastern regions. Most subzones only contain 1 to 5 businesses, with a few areas reaching 10 or more. The spatial pattern suggests that fitness centres are present in both residential neighbourhoods and near city-fringe areas, aiming to serve everyday community needs rather than concentrating solely in the central business district.

```{r}
fitness_with_zone <- st_join(biz_fitness_sf, mpsz, join = st_within)

fitness_count <- fitness_with_zone %>%
  st_drop_geometry() %>%
  group_by(SUBZONE_N) %>%
  summarise(count = n())

mpsz_fitness <- mpsz %>%
  left_join(fitness_count, by = "SUBZONE_N")

tm_shape(mpsz_fitness) +
  tm_polygons("count", palette = "Blues", title = "Fitness Businesses") +
tm_shape(biz_fitness_sf) +
  tm_dots(col = "red", size = 0.05)
```
:::

```{r}
tmap_mode("plot")
```

# 6 Point Pattern Preparation

## 6.1 Converting Boundary to Owin Object

In this step, the Master Plan 2019 Subzone Boundary (mpsz) dataset, which was originally stored as an sf object, was converted into an owin (observation window) object using the as.owin() function. The owin format is a fundamental spatial data structure in the spatstat package, required for point pattern analysis as it defines the study region (the spatial window within which points are observed and analyzed).

After conversion, we checked the class of the object to ensure it was successfully transformed, and then plotted the boundary. From the figure below, we can see the resulting map boundary of Singapore, which will serve as the analytical window for subsequent spatial point pattern analysis.

```{r}
sg_owin <- as.owin(mpsz)
```

```{r}
class(sg_owin)
```

```{r}
plot(sg_owin)
```

## 6.2 Converting Business Data to PPP Objects

In this step, the three business datasets (clinics, massage services, and fitness centres), which were initially stored as sf objects, were converted into ppp (planar point pattern) objects using the as.ppp() function from the spatstat package. The ppp format is essential for spatial point pattern analysis, as it represents each business location as a point within a defined observation window.

::: panel-tabset
# Clinics

```{r}
clinic_ppp <- as.ppp(biz_clinic_sf) %>%
  rjitter(retry = TRUE, nsim = 1, drop = TRUE)
class(clinic_ppp)
```

# Massage

```{r}
Massage_ppp <- as.ppp(biz_Massage_sf) %>%
  rjitter(retry = TRUE, nsim = 1, drop = TRUE)
class(Massage_ppp)
```

# Fitness Centres

```{r}
fitness_ppp <- as.ppp(biz_fitness_sf) %>%
  rjitter(retry = TRUE, nsim = 1, drop = TRUE)
class(fitness_ppp)
```
:::

## 6.3 Cropping Point Patterns to Study Window

After converting the business datasets into ppp objects, each point pattern was cropped to the Singapore study window (sg_owin). This ensures that only the points located within the official Singapore boundary are retained for analysis, while any points falling outside the boundary are removed.

We can see the corresponding data quantity for each item from the output results below.

::: panel-tabset
# Clinics

```{r}
clinic_ppp = clinic_ppp[sg_owin]
```

```{r}
clinic_ppp
```

# Massage

```{r}
Massage_ppp = Massage_ppp[sg_owin]
```

```{r}
Massage_ppp
```

# Fitness Centres

```{r}
fit_ppp = fitness_ppp[sg_owin]
```

```{r}
fit_ppp
```
:::

# 7 First-Order Spatial Analysis

## 7.1 Nearest Neighbour Analysis (NNA)

Nearest Neighbor Analysis (NNA) is a spatial statistics method that calculates the average distance between each point and its closest neighbor to determine if a pattern of points is clustered, dispersed, or randomly distributed.

Clark-Evans test is a specific statistical method used within NNA to quantify whether a point pattern is clustered, random, or uniformly spaced, using the Clark-Evans aggregation index (R) to describe this pattern. NNA provides a numerical value that describes the degree of clustering or regularity, and the Clark-Evans test calculates a specific index (R) for this purpose.

::: panel-tabset
# Clinics

• H₀ (Null Hypothesis): The distribution of clinics is randomly distributed.

• H₁ (Alternative Hypothesis): The distribution of clinics is not randomly distributed.

Result: R = 0.40285, p-value \< 2.2e-16. Reject H₀. The distribution of clinics is not random, with a clustered pattern across Singapore.

```{r}
clarkevans.test(clinic_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"))
```

# Massage

• H₀ (Null Hypothesis): The distribution of massage services is randomly distributed.

• H₁ (Alternative Hypothesis): The distribution of massage services is not randomly distributed.

Result: R = 0.3682, p-value \< 2.2e-16. Reject H₀. The distribution of massage services is not random, showing the strongest clustering among the three industries.

```{r}
clarkevans.test(Massage_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"))
```

# Fitness Centres

• H₀ (Null Hypothesis): The distribution of fitness centres is randomly distributed.

• H₁ (Alternative Hypothesis): The distribution of fitness centres is not randomly distributed.

Result: R = 0.41085, p-value \< 2.2e-16. Reject H₀. The distribution of fitness centres is not random, but significantly clustered.

```{r}
clarkevans.test(fit_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"))
```
:::

## 7.2 Kernel Density Estimation (KDE)

Kernel Density Estimation (KDE) is a valuable tool for visualising and analyzing first-order spatial point patterns. It is widely considered a method within Exploratory Spatial Data Analysis (ESDA) because it’s used to visualize and understand spatial data patterns by transforms discrete point data (like locations of childcare service, crime incidents or disease cases) into continuous density surfaces that reveal clusters and variations in event occurrences, without making prior assumptions about data distribution. It helps to begin understanding data distribution, identify hotspots, and explore relationships between spatial variables before performing more rigorous analysis.

## 7.2.1 Data Prepare for KDE

First, the code performs a scale transformation on the point pattern objects of the three types of points of interest (clinics, massage parlors, and fitness centers). By using rescale.ppp(..., 1000, "km"), the original spatial point data are rescaled into a coordinate system measured in kilometers, which facilitates subsequent density estimation and comparative analysis.

Next, the density() function is applied in combination with Diggle’s bandwidth selection method (bw.diggle) to conduct Kernel Density Estimation (KDE). The Gaussian kernel is chosen, and the option edge=TRUE is enabled for edge correction, in order to minimize the impact of boundary effects on the density estimation. This step produces raster objects representing the spatial intensity distributions of the three facility types.

Subsequently, the KDE results are converted from spatstat density map objects into raster objects (rast) of the terra package, and the Singapore projected coordinate system EPSG:3414 is assigned as the coordinate reference system (CRS). This ensures that the KDE results of the three facility types can be visualized, overlaid, and further analyzed under a unified spatial reference framework.

```{r}
clinic_ppp_km <- rescale.ppp(
  clinic_ppp, 1000, "km")

kde_clinic_diggle_km <- density(clinic_ppp_km,
                                sigma = bw.diggle,
                                edge = TRUE,
                                kernel = "gaussian")

massage_ppp_km <- rescale.ppp(
  Massage_ppp, 1000, "km")

kde_massage_diggle_km <- density(massage_ppp_km,
                                 sigma = bw.diggle,
                                 edge = TRUE,
                                 kernel = "gaussian")

fit_ppp_km <- rescale.ppp(fit_ppp, 1000, "km")

kde_fit_diggle_km <- density(fit_ppp_km,
                              sigma=bw.diggle,
                              edge=TRUE,
                              kernel="gaussian")

```

```{r}
kde_clinic_bw_terra <- rast(kde_clinic_diggle_km)
crs(kde_clinic_bw_terra) <- "EPSG:3414"
class(kde_clinic_bw_terra)
kde_clinic_bw_terra

kde_massage_bw_terra <- rast(kde_massage_diggle_km)
crs(kde_massage_bw_terra) <- "EPSG:3414"
class(kde_massage_bw_terra)
kde_massage_bw_terra

kde_fit_bw_terra <- rast(kde_fit_diggle_km)
crs(kde_fit_bw_terra) <- "EPSG:3414"
class(kde_fit_bw_terra)
kde_fit_bw_terra
```

### 7.2.2 Cartographic quality KDE map

Through the visualization of KDE analysis, it can be seen that the main concentration areas of the clinic, massage, and fitness industries are all located in the city center. There are also small clusters in the northern and eastern parts close to the city center. This is consistent with the analysis results in the Choropleth Maps. Therefore, we will continue to use the previously selected areas for in-depth analysis in the next step. For clinics, NOVENA and GAYLANG were chosen. For massage, OUTRAM and KALLANG were selected. For fitness, DOWNTOWN CORE and GAYLANG were chosen.

::: panel-tabset
# Clinics

```{r}
tm_shape(kde_clinic_bw_terra) + 
  tm_raster(col.scale = 
              tm_scale_continuous(
                values = "viridis"),
            col.legend = tm_legend(
            title = "Density values",
            title.size = 0.7,
            text.size = 0.7,
            bg.color = "white",
            bg.alpha = 0.7,
            position = tm_pos_in(
              "right", "bottom"),
            frame = TRUE)) +
  tm_graticules(labels.size = 0.7) +
  tm_compass() +
  tm_layout(scale = 1.0)
```

# Massage

```{r}
tm_shape(kde_massage_bw_terra) + 
  tm_raster(col.scale = 
              tm_scale_continuous(
                values = "viridis"),
            col.legend = tm_legend(
            title = "Density values",
            title.size = 0.7,
            text.size = 0.7,
            bg.color = "white",
            bg.alpha = 0.7,
            position = tm_pos_in(
              "right", "bottom"),
            frame = TRUE)) +
  tm_graticules(labels.size = 0.7) +
  tm_compass() +
  tm_layout(scale = 1.0)
```

# Fitness Centres

```{r}
tm_shape(kde_fit_bw_terra) + 
  tm_raster(col.scale = 
              tm_scale_continuous(
                values = "viridis"),
            col.legend = tm_legend(
            title = "Density values",
            title.size = 0.7,
            text.size = 0.7,
            bg.color = "white",
            bg.alpha = 0.7,
            position = tm_pos_in(
              "right", "bottom"),
            frame = TRUE)) +
  tm_graticules(labels.size = 0.7) +
  tm_compass() +
  tm_layout(scale = 1.0)
```
:::

# 8 Second-Order Spatial Analysis

## 8.1 Subset by Planning area

Extract the 6 regions from the mpsz file that correspond to the 3 industries previously selected.

```{r}
# clinic
no <- mpsz %>%
  filter(PLN_AREA_N == "NOVENA")
ge <- mpsz %>%
  filter(PLN_AREA_N == "GEYLANG")
#massage
ou <- mpsz %>%
  filter(PLN_AREA_N == "OUTRAM")
ka <- mpsz %>%
  filter(PLN_AREA_N == "KALLANG")
# fitness
dc <- mpsz %>%
  filter(PLN_AREA_N == "DOWNTOWN CORE")
# fitness also includ "GEYLANG"
```

## 8.2 Plot Planning Area Geometries

Quickly visualize the polygon geometries of the five planning areas, check whether the boundaries are correct, and whether there are any invalid geometries or coordinate issues.

```{r}
par(mfrow = c(2,3))   

# clinic
plot(st_geometry(no), main = "NOVENA")
plot(st_geometry(ge), main = "GEYLANG")

# massage
plot(st_geometry(ou), main = "OUTRAM")
plot(st_geometry(ka), main = "KALLANG")

# fitness
plot(st_geometry(dc), main = "DOWNTOWN CORE")
```

## 8.3 Convert Planning Areas to Windows (owin)

Convert each planning area polygon into an observation window owin in spatstat, which will serve as the research boundary for subsequent point pattern clipping and analysis.

```{r}
# clinic
no_owin = as.owin(no)
ge_owin = as.owin(ge)

# massage
ou_owin = as.owin(ou)
ka_owin = as.owin(ka)

# fitness
dc_owin = as.owin(dc)
```

## 8.4 Point Patterns per Planning Area (Clinics / Massage / Fitness)

The point pattern data were clipped to their corresponding planning area windows to retain only events within each boundary. The patterns were then rescaled from meters to kilometers using rescale.ppp(..., 1000, "km"), and visualized with plot(unmark(...)) to provide a quick view of the distribution, allowing an intuitive comparison of density and spatial structure across regions.

::: panel-tabset
# Clinics

In Novena, clinics demonstrate a strong core clustering effect, with almost all facilities located around the medical hub and areas with good transport accessibility. This aligns with Novena’s status as one of Singapore’s major medical hubs, showing that the clinic industry here is highly concentrated under the influence of regional healthcare resources.

In Geylang, the number of clinics is relatively high, and their distribution is more dispersed, covering almost the entire area. This layout suggests that the clinic industry in Geylang primarily aims to meet community-level healthcare needs, emphasizing accessibility and balanced regional distribution. At the same time, it may also be associated with the presence of the sex trade in Geylang. This contrasts with Novena’s positioning as a specialized medical hub.

```{r}
clinic_no_ppp  = clinic_ppp[no_owin]
clinic_ge_ppp  = clinic_ppp[ge_owin]

clinic_no_ppp.km  = rescale.ppp(clinic_no_ppp, 1000, "km")
clinic_ge_ppp.km  = rescale.ppp(clinic_ge_ppp, 1000, "km")

par(mfrow = c(1,2), cex = 0.6)
plot(unmark(clinic_no_ppp.km), main = "NOVENA (Clinic)")
plot(unmark(clinic_ge_ppp.km), main = "GEYLANG (Clinic)")
```

# Massage

In Outram, massage service outlets are relatively numerous and evenly distributed, with a focus on areas of strong transport connectivity and commercial activity. This distribution is consistent with Outram’s role as a district with medical, commercial, and leisure functions, reflecting the positioning of the massage industry here as primarily leisure and wellness services.

In Kallang, the number of massage outlets is also considerable, but they are more heavily concentrated in the southern boundary area, possibly due to proximity to transportation hubs and commercial centers. This clustering indicates that the massage industry in Kallang tends to align with transport and consumption nodes, forming localized, high-density service zones.

```{r}
massage_ou_ppp  = Massage_ppp[ou_owin]
massage_ka_ppp  = Massage_ppp[ka_owin]

massage_ou_ppp.km  = rescale.ppp(massage_ou_ppp, 1000, "km")
massage_ka_ppp.km  = rescale.ppp(massage_ka_ppp, 1000, "km")

par(mfrow = c(1,2), cex = 0.6)
plot(unmark(massage_ou_ppp.km), main = "OUTRAM (Massage)")
plot(unmark(massage_ka_ppp.km), main = "KALLANG (Massage)")

```

# Fitness Centres

In Downtown Core, fitness centers exhibit a clear clustering pattern, particularly concentrated in the southern and western areas. This aligns closely with the district’s role as a commercial and office hub. The high density of fitness centers in locations with a large working population reflects the strong correlation between fitness demand and workplace population density.

In contrast, in Geylang, fitness centers are more dispersed, with points scattered across the area and only small-scale clusters in certain spots. This pattern suggests that in Geylang, fitness centers primarily serve the residential population and community-based consumption needs, rather than relying heavily on concentrated demand from business districts.

```{r}
fit_dc_ppp  = fitness_ppp[dc_owin]
fit_ge_ppp  = fitness_ppp[ge_owin]

fit_dc_ppp.km  = rescale.ppp(fit_dc_ppp, 1000, "km")
fit_ge_ppp.km  = rescale.ppp(fit_ge_ppp, 1000, "km")
par(mfrow = c(1,2), cex = 0.6)
plot(unmark(fit_dc_ppp.km), main = "DOWNTOWN CORE (Fitness)")
plot(unmark(fit_ge_ppp.km), main = "GEYLANG (Fitness)")
```
:::

## 8.5 Setting Random Seed

The random seed is fixed with set.seed(1234) to ensure reproducibility of stochastic spatial simulations and statistical tests, so that the results can be replicated consistently.

```{r}
set.seed(1234)
```

## 8.6 Second-Order Spatial Analysis For Clinic

The second-order spatial analysis of clinics in both Geylang and Novena indicates significant clustering, as the observed K(d) and L(d) values are consistently higher than the CSR benchmark. Thus, the null hypothesis of random distribution is rejected in both cases. However, the nature of clustering differs between the two areas.

Although neither Geylang nor Novena is part of the city’s central business district, their clinic distributions show distinct spatial patterns. Novena exhibits a highly concentrated core clustering, reflecting its role as a major healthcare hub where clinics agglomerate around specialized medical facilities and transport nodes. By contrast, Geylang shows a more dispersed form of clustering, with clinics spread relatively evenly across the area. This suggests that while clustering exists, it is community-oriented and designed to ensure accessibility for local residents, rather than forming a tightly bounded medical core.

::::: panel-tabset
# Novena

Ho = The distribution of Clinic at NOVENA are randomly distributed.

H1= The distribution of Clinic at NOVENA are not randomly distributed.

Based on the calculation of the K-function and L-function, both K(d) and L(d) are significantly higher than CSR. Therefore, we reject H_0 and accept H_1, indicating that clinic establishments in the Novena area are significantly clustered.

::: panel-tabset
## K-function

```{r}
K_clinic_no = Kest(clinic_no_ppp, correction = "Ripley")
plot(K_clinic_no, . -r ~ r, ylab= "K(d)-r", xlab = "d(m)")
```

```{r}
K_clinic_no.csr <- envelope(clinic_no_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(K_clinic_no.csr, . - r ~ r, xlab="d", ylab="K(d)-r")
```

## L-function

```{r}
L_clinic_no = Lest(clinic_no_ppp, correction = "Ripley")
plot(L_clinic_no, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)")
```

```{r}
L_clinic_no.csr <- envelope(clinic_no_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(L_clinic_no.csr, . - r ~ r, xlab="d", ylab="L(d)-r")
```
:::

# Geylang

Ho = The distribution of Clinic at Geylang are randomly distributed.

H1= The distribution of Clinic at Geylang are not randomly distributed.

Based on the calculation of the K-function and L-function, both K(d) and L(d) are significantly higher than CSR. Therefore, we reject H_0 and accept H_1, indicating that clinic establishments in the Geylang area are significantly clustered.

::: panel-tabset
## K-function

```{r}
K_clinic_ge = Kest(clinic_ge_ppp, correction = "Ripley")
plot(K_clinic_ge, . -r ~ r, ylab= "K(d)-r", xlab = "d(m)")
```

```{r}
K_clinic_ge.csr <- envelope(clinic_ge_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(K_clinic_ge.csr, . - r ~ r, xlab="d", ylab="K(d)-r")
```

## L-function

```{r}
L_clinic_ge = Lest(clinic_ge_ppp, correction = "Ripley")
plot(L_clinic_ge, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)")
```

```{r}
L_clinic_ge.csr <- envelope(clinic_ge_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(L_clinic_ge.csr, . - r ~ r, xlab="d", ylab="L(d)-r")
```
:::
:::::

## 8.7 Second-Order Spatial Analysis For Massage

The second-order spatial analysis shows that both Outram and Kallang exhibit significant clustering of massage establishments, as indicated by K(d) and L(d) values exceeding the CSR benchmark. However, the nature and intensity of clustering differ across the two regions.

In Outram, massage outlets display a moderately clustered yet relatively balanced distribution, with establishments concentrated around central accessible areas but also dispersed into smaller sub-clusters across the region. This suggests that the massage industry in Outram is integrated into the district’s mixed-use character, combining healthcare, leisure, and commercial activities.

By contrast, in Kallang, clustering is much more pronounced and localized, with a dense concentration of establishments in the southern part of the district, near major transport nodes and commercial corridors. Beyond this hotspot, the distribution becomes more scattered. This indicates that in Kallang, the massage industry is more dependent on high-traffic and high-consumption nodes, forming a strong local cluster rather than a balanced spread.

::::: panel-tabset
# Outram

Ho = The distribution of Massage at Outram are randomly distributed.

H1= The distribution of Massage at Outram are not randomly distributed.

Based on the calculation of the K-function and L-function, both K(d) and L(d) are significantly higher than CSR. Therefore, we reject H_0 and accept H_1, indicating that massage establishments in the Outram area are significantly clustered.

::: panel-tabset
## K-function

```{r}
K_massage_ou = Kest(massage_ou_ppp, correction = "Ripley")
plot(K_massage_ou, . -r ~ r, ylab= "K(d)-r", xlab = "d(m)")
```

```{r}
K_massage_ou.csr <- envelope(massage_ou_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(K_massage_ou.csr, . - r ~ r, xlab="d", ylab="K(d)-r")
```

## L-function

```{r}
L_massage_ou = Lest(massage_ou_ppp, correction = "Ripley")
plot(L_massage_ou, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)")
```

```{r}
L_massage_ou.csr <- envelope(massage_ou_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(L_massage_ou.csr, . - r ~ r, xlab="d", ylab="L(d)-r")
```
:::

# Kallang

Ho = The distribution of Massage at Outram are randomly distributed.

H1= The distribution of Massage at Outram are not randomly distributed.

Based on the calculation of the K-function and L-function, both K(d) and L(d) are significantly higher than CSR. Therefore, we reject H_0 and accept H_1, indicating that massage establishments in the Kallang area are significantly clustered.

::: panel-tabset
## K-function

```{r}
K_massage_ka = Kest(massage_ka_ppp, correction = "Ripley")
plot(K_massage_ka, . -r ~ r, ylab= "K(d)-r", xlab = "d(m)")
```

```{r}
K_massage_ka.csr <- envelope(massage_ka_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(K_massage_ka.csr, . - r ~ r, xlab="d", ylab="K(d)-r")
```

## L-function

```{r}
L_massage_ka = Lest(massage_ka_ppp, correction = "Ripley")
plot(L_massage_ka, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)")
```

```{r}
L_massage_ka.csr <- envelope(massage_ka_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(L_massage_ka.csr, . - r ~ r, xlab="d", ylab="L(d)-r")
```
:::
:::::

## 8.8 Second-Order Spatial Analysis For Fitness

Based on the results of second-order spatial analysis, the distribution of fitness centers in both Geylang and Downtown Core significantly deviates from complete spatial randomness, showing clear clustering patterns. However, the clustering characteristics differ between the two areas.

In Downtown Core, fitness centers are highly concentrated in the southern and southwestern parts of the district, forming a compact core cluster. This aligns with the area’s role as a commercial and office hub with a dense working population. The distribution pattern of fitness centers here is strongly linked to workplace demand.

In contrast, in Geylang, fitness centers also exhibit significant clustering, but the overall distribution is more dispersed. Facilities are spread across the area, with only small-scale clusters forming in certain locations. This indicates that fitness centers in Geylang are more oriented toward serving the residential population and everyday community needs, rather than being driven primarily by business or workplace demand.

::::: panel-tabset
# Downtown Core

Ho = The distribution of Fitness at Downtown Core are randomly distributed.

H1 = The distribution of Fitness at Downtown Core are not randomly distributed.

Based on the calculation of the K-function and L-function, both K(d) and L(d) are significantly higher than CSR. Therefore, we reject H_0 and accept H_1, indicating that fitness establishments in the Downtown Core area are significantly clustered.

::: panel-tabset
## K-function

```{r}
K_fit_dc = Kest(fit_dc_ppp, correction = "Ripley")
plot(K_fit_dc, . -r ~ r, ylab= "K(d)-r", xlab = "d(m)")
```

```{r}
K_fit_dc.csr <- envelope(fit_dc_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(K_fit_dc.csr, . - r ~ r, xlab="d", ylab="K(d)-r")
```

## L-function

```{r}
L_fit_dc = Lest(fit_dc_ppp, correction = "Ripley")
plot(L_fit_dc, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)")
```

```{r}
L_fit_dc.csr <- envelope(fit_dc_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(L_fit_dc.csr, . - r ~ r, xlab="d", ylab="L(d)-r")
```
:::

# Geylang

Ho = The distribution of Fitness at Geylang are randomly distributed.

H1= The distribution of Fitness at Geylang are not randomly distributed.

Based on the calculation of the K-function and L-function, both K(d) and L(d) are significantly higher than CSR. Therefore, we reject H_0 and accept H_1, indicating that fitness establishments in the Gelang area are significantly clustered.

::: panel-tabset
## K-function

```{r}
K_fit_ge = Kest(fit_ge_ppp, correction = "Ripley")
plot(K_fit_ge, . -r ~ r, ylab= "K(d)-r", xlab = "d(m)")
```

```{r}
K_fit_ge.csr <- envelope(fit_ge_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(K_fit_ge.csr, . - r ~ r, xlab="d", ylab="K(d)-r")
```

## L-function

```{r}
L_fit_ge = Lest(fit_ge_ppp, correction = "Ripley")
plot(L_fit_ge, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)")
```

```{r}
L_fit_ge.csr <- envelope(fit_ge_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(L_fit_ge.csr, . - r ~ r, xlab="d", ylab="L(d)-r")
```
:::
:::::

## 9 Spatio-temporal Point Patterns Analysis: stpp methods

### 9.1 Data Prepare for stpp

In this step, the three business datasets (clinics, massage, and fitness centres) were first converted into sf objects with valid spatial coordinates (EPSG:3414). Each dataset was then spatially joined with the URA Master Plan 2019 planning area boundaries (mpsz) to append the corresponding planning area attributes.

From these enriched datasets, six sub-areas of interest were selected for further analysis:

• Clinics: Novena and Geylang

• Massage: Outram and Kallang

• Fitness centres: Downtown Core and Geylang

The spatial coordinates (x, y) and registration dates (t) of each subset were extracted and assembled into data frames, which were then transformed into 3D spatio-temporal point pattern objects using as.3dpoints(). Finally, the spatio-temporal K-function (STIKhat) was applied to each subset to measure clustering behaviour over both spatial and temporal dimensions.

```{r}

biz_clinic_sf <- st_as_sf(
  biz_clinic %>% filter(!is.na(X), !is.na(Y)),
  coords = c("X","Y"), crs = 3414
)

biz_massage_sf <- st_as_sf(
  biz_Massage %>% filter(!is.na(X), !is.na(Y)),
  coords = c("X","Y"), crs = 3414
)

biz_fitness_sf <- st_as_sf(
  biz_fitness %>% filter(!is.na(X), !is.na(Y)),
  coords = c("X","Y"), crs = 3414
)

biz_clinic_sf <- st_join(biz_clinic_sf, mpsz, join = st_within)
biz_massage_sf <- st_join(biz_massage_sf, mpsz, join = st_within)
biz_fitness_sf <- st_join(biz_fitness_sf, mpsz, join = st_within)

clinic_no_sf <- biz_clinic_sf %>%
  filter(PLN_AREA_N == "NOVENA")
clinic_ge_sf <- biz_clinic_sf %>%
  filter(PLN_AREA_N == "GEYLANG")
massage_ou_sf <- biz_massage_sf %>%
  filter(PLN_AREA_N == "OUTRAM")
massage_ka_sf <- biz_massage_sf %>%
  filter(PLN_AREA_N == "KALLANG")
fitness_dc_sf <- biz_fitness_sf %>%
  filter(PLN_AREA_N == "DOWNTOWN CORE")
fitness_ge_sf <- biz_fitness_sf %>%
  filter(PLN_AREA_N == "GEYLANG")

coords_c_no <- st_coordinates(clinic_no_sf)   # Novena
coords_c_ge <- st_coordinates(clinic_ge_sf)   # Geylang
coords_m_ou <- st_coordinates(massage_ou_sf)  # Outram
coords_m_ka <- st_coordinates(massage_ka_sf)  # Kallang
coords_f_dc <- st_coordinates(fitness_dc_sf)  # Downtown Core
coords_f_ge <- st_coordinates(fitness_ge_sf)  # Geylang


# --- Clinics: Novena ---
clinic_no_df <- data.frame(
  x = coords_c_no[, 1],  
  y = coords_c_no[, 2],
  t = clinic_no_sf$date
)
clinic_no_stpp <- as.3dpoints(clinic_no_df)
kbb_clinic_no <- STIKhat(clinic_no_stpp)

# --- Clinics: Geylang ---
clinic_ge_df <- data.frame(
  x = coords_c_ge[, 1],  
  y = coords_c_ge[, 2],
  t = clinic_ge_sf$date
)
clinic_ge_stpp <- as.3dpoints(clinic_ge_df)
kbb_clinic_ge <- STIKhat(clinic_ge_stpp)

# --- Massage: Outram ---
massage_ou_df <- data.frame(
  x = coords_m_ou[, 1],  
  y = coords_m_ou[, 2],
  t = massage_ou_sf$date
)
massage_ou_stpp <- as.3dpoints(massage_ou_df)
kbb_massage_ou <- STIKhat(massage_ou_stpp)

# --- Massage: Kallang ---
massage_ka_df <- data.frame(
  x = coords_m_ka[, 1],  
  y = coords_m_ka[, 2],
  t = massage_ka_sf$date
)
massage_ka_stpp <- as.3dpoints(massage_ka_df)
kbb_massage_ka <- STIKhat(massage_ka_stpp)

# --- Fitness: Downtown Core ---
fitness_dc_df <- data.frame(
  x = coords_f_dc[, 1],  
  y = coords_f_dc[, 2],
  t = fitness_dc_sf$date
)
fitness_dc_stpp <- as.3dpoints(fitness_dc_df)
kbb_fitness_dc <- STIKhat(fitness_dc_stpp)

# --- Fitness: Geylang ---
fitness_ge_df <- data.frame(
  x = coords_f_ge[, 1],  
  y = coords_f_ge[, 2],
  t = fitness_ge_sf$date
)
fitness_ge_stpp <- as.3dpoints(fitness_ge_df)
kbb_fitness_ge <- STIKhat(fitness_ge_stpp)

```

### 9.2 Spatio-temporal Point Patterns Analysis

::: panel-tabset
# clinic

Novena: The contour plot shows moderate clustering with flatter, lighter lines at short spatial and temporal distances. Clustering exists but is localized and less intense, consistent with Novena’s role as a structured medical hub where clinics cluster around a core but without strong temporal bursts.

Geylang: Steeper and darker contour lines indicate stronger clustering across larger spatial and temporal scales. Clinics in Geylang are not only spatially close but also appear more synchronously in time, reflecting community-driven demand and dynamic bursts of activity.

Comparison: Both areas show clustering, but Novena’s is moderate and core-focused, while Geylang’s is stronger and more widespread, tied to broader local healthcare demand.

```{r}
par(mfrow = c(1, 2))  
plotK(kbb_clinic_no, main = "Clinic - Novena")
plotK(kbb_clinic_ge, main = "Clinic - Geylang")
par(mfrow = c(1, 1))  
```

# massage

Outram shows weaker clustering, with flatter and lighter contour lines concentrated in the lower ranges of spatial and temporal distances. This indicates that while clustering occurs, it is more localized and less intense. Outram’s clustering reflects steady demand tied to its mixed medical, leisure, and commercial context, but without strong temporal bursts.

Kallang, in contrast, demonstrates much stronger clustering, with steeper and darker contours that extend over broader spatial and temporal scales. This suggests that massage businesses tend to open closer together both in time and in space, forming stronger synchronized clusters. The pattern aligns with Kallang’s proximity to transport hubs and entertainment areas, which may drive denser and more simultaneous establishment activity.

Comparison: Outram’s clustering is moderate and localized, while Kallang shows more intense and widespread clustering, reflecting higher spatial-temporal concentration of business activity.

```{r}
par(mfrow = c(1, 2))  
plotK(kbb_massage_ou, main = "Massage - Outram")
plotK(kbb_massage_ka, main = "Massage - Kallang")
par(mfrow = c(1, 1))  
```

# fitness

Downtown Core: The contour lines rise steeply and darken at larger spatial and temporal distances, indicating that clustering is not only strong in the short term (centres opening close in both time and space) but also sustained over longer periods and wider spatial ranges. This reflects persistent demand from a dense working population, producing both immediate bursts and long-term accumulation of fitness businesses.

Geylang: The contours are flatter and lighter, showing clustering that is present but weaker. Clustering occurs mainly at short spatial and temporal scales, suggesting more localized bursts of activity. Over longer periods, the clustering effect is less pronounced, pointing to a community-driven but less synchronized growth pattern.

Comparison: While both areas demonstrate spatio-temporal clustering, Downtown Core shows both short-term bursts and long-term sustained clustering, whereas Geylang’s clustering is more localized and concentrated within shorter time spans.

```{r}
par(mfrow = c(1, 2))   
plotK(kbb_fitness_dc, main = "Fitness - Downtown Core")
plotK(kbb_fitness_ge, main = "Fitness - Geylang")
par(mfrow = c(1, 1))  
```
:::

# 10 Conclusion

1.  To avoid homogeneous competition, the Massage industry in Outram and Kallang shows high-density agglomeration with service models dominated by foot reflexology and traditional massage, leading to rent escalation and neighborhood disturbance risks. It is recommended to establish capacity caps along metro stations and main streets, introduce minimum spacing and street-level upper limits, and encourage upgrading from single-function massage outlets to “wellness complexes” that integrate physiotherapy and sports rehabilitation. Such integration with fitness centers and clinics can create complementary service ecosystems, simultaneously enhancing convenience for residents while alleviating the crowding-out effect of single-industry clustering.

2.  Regarding healthcare provision, although Clinics are distributed island-wide, their concentration is stronger in the east and central zones (such as Novena and Geylang), while the northwestern and northern peripheries remain relatively sparse. To close this gap, healthcare and other essential public services should expand coverage by leveraging transport hubs as anchors for new medical clusters. In areas such as Woodlands and Jurong West, the addition of family clinics and extended-hours outpatient facilities can improve accessibility and reduce the need for cross-district travel, particularly during night-time hours.

3.  The Fitness industry exhibits intense clustering in Downtown Core, characterized by short-term bursts and long-term persistence. This produces pressure on transport systems and public spaces during peak commuting periods, while excessive competition among enterprises further intensifies the issue. To address this, demand should be gradually dispersed toward communities through improved connectivity and expansion of chain outlets beyond the CBD. Outside the CBD, where density is low and distribution is fragmented, the “15-minute fitness circle” should be promoted by introducing small-scale gyms centered on residential neighborhoods, combined with park greenways and walking networks. This would provide affordable, community-level fitness opportunities and prevent excessive reliance on the CBD.

4.  At the regulatory and planning level, spatio-temporal point pattern (STPP) analysis reveals simultaneous “opening pulses” in several districts within the same month. It is recommended to pilot “quarterly rolling quotas with time-window approvals” in Outram, Kallang, and Downtown Core, with thresholds calibrated against noise, fire safety, pedestrian capacity, and waste management. In hotspot streets, a “pre-assessment before expansion” trial permit system should be applied, linking licensing directly to compliance scores. Establishments with low compliance should only receive short-term permits and be required to implement corrective measures. This would curb unrestrained expansion and mitigate the risk of adverse selection in the business ecosystem.

# 11 Reference

-   Kam, T. S. (2024). *R for Geospatial Data Science and Analytics* <https://r4gdsa.netlify.app>
