---
title: "Take-home Exercise 1"
author: "Li Yuquan"
description: ""
date: "September 15, 2025"
date-modified: "last-modified"
format: html
editor: visual
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# 1 Overview

## 1.1 Objectives

The location choices of businesses play a pivotal role in shaping urban development, influencing economic growth, job creation, and the provision of resources and services. Strategic siting not only creates competitive advantages and supports efficient supply chains, but also contributes to a city’s identity through specialized districts and accessible infrastructure.

Studying the spatio-temporal patterns of new businesses is therefore essential, as it highlights how economic activity and urban forms evolve over time. Such analysis enables urban planners to anticipate growth, identify clusters and trends, improve resource allocation, and design more sustainable and livable cities.

In this exercise, the focus is on new business entities established in Singapore during the first six months of 2025. The objectives are:

• To analyze the “where” and “when” of new business establishments, uncovering spatial clusters, temporal trends, and development hotspots.

• To explore complex interaction patterns beyond simple density measures, such as the clustering of related firms, path-dependency of business locations, and structural changes in commercial areas.

## 1.2 The Task

1.  Download ACRA (Accounting and Corporate Regulatory Authority) Information on Corporate Entities datasets from data.gov.sg.

2.  Select at least three business types by referring to [2-digit or 3-digit Singapore Standard Industrial Classification (SSIS) 2020](https://classification.codes/classifications/industry/ssic-singapore). Prepare and tidy the data by applying appropriate data wrangling method.

3.  Extract the selected business entities registered between 1st January 2024 to 30th June 2025.

4.  Geocode the selected records and convert them into simple point features.

5.  Compute spatial and spatio-temporal KDE of the selected business types.

6.  Perform second order spatial and spatio-temporal analysis on the selected business types.

7.  Describe and discuss the analysis results obtained from Step 5 and 6.

8.  For each geovisualisation prepared, write a short report of not more than 150 words describing the spatio-temporal patterns revealed by the geovisualisation and statistical graphics.

9.  With reference to the analysis results above, discuss how these findings can be used to support urban land use planning and management. (Not more than 500 words)

# 1.3 R packages

| **Package** | **Description** |
|------------------------------------|------------------------------------|
| sf | Simple Features for handling and analyzing spatial vector data. |
| spNetwork | Network-based spatial analysis, e.g. kernel density along road networks. |
| spatstat | Spatial point pattern analysis (ppp, owin, G/F/K/L functions, KDE). |
| tmap | Thematic maps creation (static and interactive). |
| tidyverse | Data science suite (dplyr, ggplot2, readr, purrr, etc.). |
| httr | Tools for working with HTTP to access web APIs. |
| rvest | Web scraping and HTML parsing. |
| terra | Spatial raster data handling and analysis. |
| lwgeom | Geometry operations, fixing invalid shapes (wraps GEOS). |

```{r}
pacman::p_load(sf, spNetwork, spatstat,tmap, tidyverse, httr, rvest, terra, lwgeom, stpp)
```

# 1.4 The Data

| **Dataset Name** | **Description** | **Format** | **Source** |
|------------------|------------------|------------------|------------------|
| ACRA Information on Corporate Entities ('A-Z-Others') | Business registration information (clinics, massage, fitness filtered by SSIC) | CSV | data.gov.sg |
| OneMap Geocoding Results | Geocoded coordinates (X, Y) for postal codes | API | [OneMap API (Singapore Land Authority)](https://onemap.gov.sg/api/common/elastic/search) |
| MasterPlan2019SubzoneBoundaryNoSeaKML | Planning region and subzone polygon boundaries | KML | URA (Urban Redevelopment Authority) |
| Singapore Standard Industrial Classification (SSIS) 2020 | SSIS Code Matching Table | Excel | [2-digit or 3-digit Singapore Standard Industrial Classification (SSIS) 2020](https://classification.codes/classifications/industry/ssic-singapore) |

# 4 Data Import and Wrangling

## 4.1 Importing ACRA data

Import ACRA data to R studio, combin all 27 file togeter and save it as rds file.

```{r}

folder_path <- "data/aspatial"
file_list <- list.files(path = folder_path,
                        pattern = "^ACRA*.*\\.csv$",
                        full.names = TRUE)

acra_data <- file_list %>%
  map_dfr(read_csv)
```

```{r}
write_rds(acra_data,
          "data/rds/acra_data.rds")

```

```{r}
summary(acra_data) 
```

```{r}
glimpse(acra_data)
```

## 4.2 Importing MasterPlan2019SubzoneBoundaryNoSea data

The Master Plan 2019 Subzone Boundary (KML) file was imported, re-projected to Singapore’s SVY21 coordinate system (EPSG:3414), and saved as an RDS file for later mapping and analysis.

```{r}

mpsz <- st_read("data/MasterPlan2019SubzoneBoundaryNoSeaKML.kml")
mpsz <- mpsz %>% st_transform(crs=3414)

```

```{r}
write_rds(mpsz, "data/rds/mpsz.rds")
mpsz <- read_rds("data/rds/mpsz.rds")
```

```{r}
summary(mpsz) 
```

```{r}
glimpse(mpsz)
```

## 4.3 Tidying ACRA data

The ACRA dataset was filtered to keep only four SSIC codes representing three industry types: Clinics (86201, 86202), Massage services (96024), and Fitness centres (93111). The data was cleaned (date fields, year/month extraction, postal code formatting) and then split into three subsets: biz_clinic, biz_Massage, and biz_fitness.

| **Industry Type** | **SSIC Code(s)** | **Description** |
|-------------------|-------------------|----------------------------------|
| Clinics | 86201, 86202 | Clinics and other general medical services (Western & Non-Western) |
| Massage Services | 96024 | Massage parlours and foot reflexology |
| Fitness Centres | 93111 | Fitness centres and gymnasiums |

```{r}
ssic_vec <- c("86201", "86202", "96024", "93111")

acra_core <- acra_data %>%
  select(1:24) %>% 
  mutate(
    primary_ssic_code = as.character(primary_ssic_code),
    date  = as.Date(registration_incorporation_date),
    YEAR  = year(date),
    MONTH_NUM  = month(date),
    MONTH_ABBR = month(date, label = TRUE, abbr = TRUE)
  ) %>%
  filter(
    date >= as.Date("2024-01-01"),
    date <= as.Date("2025-06-30"),
    primary_ssic_code %in% ssic_vec
  ) %>%
  mutate(
    postal_code = str_pad(postal_code, width = 6, side = "left", pad = "0")
  )


biz_clinic  <- acra_core %>% filter(primary_ssic_code %in% c("86201", "86202"))
biz_Massage  <- acra_core %>% filter(primary_ssic_code %in% c("96024"))
biz_fitness <- acra_core %>% filter(primary_ssic_code == "93111")
```

## 4.4 Geocoding

The postal codes from the three business subsets (clinics, massage, fitness centres) were extracted and cleaned to remove duplicates and missing values. These unique postal codes were then sent to the OneMap API for geocoding. For each code, the API returned geographic coordinates (X, Y) and address details if found. The results were stored in a found dataset, while unmatched postal codes were recorded separately in a not_found dataset.

```{r}
postcodes <- c(biz_clinic$postal_code,
               biz_Massage$postal_code,
               biz_fitness$postal_code) %>%
  unique() %>%
  na.omit()

url <- "https://onemap.gov.sg/api/common/elastic/search"

found     <- data.frame()
not_found <- data.frame(postcode = character())

for (pc in postcodes) {
  query <- list(
    searchVal = pc,
    returnGeom = "Y",
    getAddrDetails = "Y",
    pageNum = "1"
  )
  res  <- GET(url, query = query)
  json <- content(res)
  
  if (!is.null(json$found) && json$found != 0) {
    df <- as.data.frame(json$results, stringsAsFactors = FALSE)
    df$input_postcode <- pc
    found <- bind_rows(found, df)
  } else {
    not_found <- bind_rows(not_found, data.frame(postcode = pc))
  }
 
}
```

## 4.5 Tidying the geocoded data

The geocoding results were cleaned by selecting only the first 10 relevant fields from the found dataset to keep essential location attributes.

```{r}
found <- found %>%
  select(1:10)
```

## 4.6 Appending the location information

The cleaned geocoded results were joined back to the three business datasets (biz_clinic, biz_Massage, biz_fitness) using postal code as the key, enriching each business record with its spatial coordinates and address details. The updated datasets were then saved as RDS files for later use.

```{r}

biz_clinic  <- biz_clinic  %>% left_join(found, by = c("postal_code" = "POSTAL"))
biz_Massage  <- biz_Massage  %>% left_join(found, by = c("postal_code" = "POSTAL"))
biz_fitness <- biz_fitness %>% left_join(found, by = c("postal_code" = "POSTAL"))
```

```{r}
write_rds(biz_clinic, "data/rds/biz_clinic.rds")
write_rds(biz_Massage, "data/rds/biz_Massage.rds")
write_rds(biz_fitness, "data/rds/biz_fitness.rds")


```

## 4.7 Converting into SF data frame

Each business dataset was transformed into an sf (simple features) object (biz_clinic_sf, biz_Massage_sf, biz_fitness_sf), keeping only records with valid coordinates (X, Y). The coordinate system was set to EPSG:3414 (SVY21 / Singapore TM), enabling spatial mapping and analysis.

```{r}
biz_clinic_sf <- st_as_sf(
  biz_clinic %>% filter(!is.na(X), !is.na(Y)),
  coords = c("X","Y"), crs = 3414
)

biz_Massage_sf <- st_as_sf(
  biz_Massage %>% filter(!is.na(X), !is.na(Y)),
  coords = c("X","Y"), crs = 3414
)

biz_fitness_sf <- st_as_sf(
  biz_fitness %>% filter(!is.na(X), !is.na(Y)),
  coords = c("X","Y"), crs = 3414
)
```

### 4.7.1 Glimpse for 3 industry

By performing the Glimpse operation on the data sets of the three industries, we can determine the amount of data contained in each one by looking at the number of rows.

::: panel-tabset
# Clinics

```{r}
glimpse(biz_clinic_sf)
```

# Massage

```{r}
glimpse(biz_Massage_sf)
```

# Fitness Centres

```{r}
glimpse(biz_fitness_sf)
```
:::

## 4.8 Extracting Attributes from KML Description

The imported Master Plan 2019 Subzone Boundary (mpsz) dataset contained key attributes (e.g., region, planning area, subzone name, subzone code) embedded inside the Description field of the KML file. A custom function extract_kml_field() was created to parse the HTML structure and extract these values. The extracted fields (REGION_N, PLN_AREA_N, SUBZONE_N, SUBZONE_C) were then added as new columns to the dataset. Finally, redundant fields (Name, Description) were removed, and the geometry column was relocated to the end for a cleaner structure.

```{r}
extract_kml_field <- function(html_text, field_name) {
  if (is.na(html_text) || html_text == "") return(NA_character_)
  
  page <- read_html(html_text)
  rows <- page %>% html_elements("tr")
  
  value <- rows %>%
    keep(~ html_text2(html_element(.x, "th")) == field_name) %>%
    html_element("td") %>%
    html_text2()
  
  if (length(value) == 0) NA_character_ else value
}

mpsz <- mpsz %>%
  mutate(
    REGION_N = map_chr(Description, extract_kml_field, "REGION_N"),
    PLN_AREA_N = map_chr(Description, extract_kml_field, "PLN_AREA_N"),
    SUBZONE_N = map_chr(Description, extract_kml_field, "SUBZONE_N"),
    SUBZONE_C = map_chr(Description, extract_kml_field, "SUBZONE_C")
  ) %>%
  select(-Name, -Description) %>%
  relocate(geometry, .after = last_col())
```

```{r}
glimpse(mpsz)
```

# 5 Exploratory Visualization

In this section, I will conduct a preliminary visualization and exploratory analysis on the processed data.

## 5.1 Visualising the distribution

::: panel-tabset
# Clinics

The number of clinic registrations in 2024 generally remained between 20 and 35 per month, with a noticeable peak in the middle of the year (May–July), reaching or exceeding 35. In the first half of 2025, the trend remained strong, with registrations in May climbing above 40, the highest value during the observation period. This suggests that the clinic industry exhibits clear seasonal concentration in registrations, with stronger growth momentum in 2025.

```{r}
# Clinics
ggplot(biz_clinic, aes(x = MONTH_ABBR)) +
  geom_bar() +
  facet_wrap(~ YEAR, ncol = 2) +
  labs(title = "Monthly Count of New Clinics, 2024–2025",
       x = "Month", y = "Count")

```

# Massage

The number of massage-related business registrations in 2024 was relatively balanced across months, mostly between 20 and 30, with little fluctuation. In the first half of 2025, registrations continued at a similar level, with small peaks above 30 in January and March 2024 and in May 2025. Overall, this industry demonstrates a stable and evenly distributed pattern without obvious seasonal fluctuations.

```{r}
# Massage
ggplot(biz_Massage, aes(x = MONTH_ABBR)) +
  geom_bar() +
  facet_wrap(~ YEAR, ncol = 2) +
  labs(title = "Monthly Count of New Massage, 2024–2025",
       x = "Month", y = "Count")

```

# Fitness Centres

The number of fitness centre registrations fluctuated widely during 2024–2025. In May 2025, a clear peak was observed with 40 new businesses, while in most other months the number hovered between 15 and 25, showing significant variation. This indicates that the fitness industry experiences instability in the number of new company registrations.

```{r}
# Fitness centres
ggplot(biz_fitness, aes(x = MONTH_ABBR)) +
  geom_bar() +
  facet_wrap(~ YEAR, ncol = 2) +
  labs(title = "Monthly Count of New Fitness Centres , 2024–2025",
       x = "Month", y = "Count")
```
:::

## 5.2 Spatial Distribution of Businesses

```{r}
tmap_mode("view")
```

::: panel-tabset
# Clinics

The distribution of clinics is relatively balanced, covering the entire island, from the city centre to outer residential districts, though overall slightly more in the east than in the west. Compared to massage and fitness centres, clinics are closer to a public service function, with wider coverage to ensure residents in different communities have access to basic healthcare. As a result, evenly spread blue dots can also be observed in western, northern, and eastern neighbourhoods, highlighting the ubiquity and accessibility of clinics.

```{r}
tm_shape(biz_clinic_sf) +
  tm_dots(col = "blue", alpha = 0.7, size = 0.06) +
  tm_view(set.view = c(103.851959, 1.290270, 11)) +
  tm_layout(title = "Locations of New Clinics")
```

# Massage

Massage businesses are more concentrated within the city area, especially in Orchard Road, Bugis, and the central-southern regions, where red dots are densely packed. This pattern is consistent with the industry’s reliance on areas with high foot traffic and active consumer spending. In the outer areas, the points are relatively fewer, showing that massage services mainly target the city centre and certain prosperous residential zones, with a strong dependence on commercial demand.

```{r}
tm_shape(biz_Massage_sf) +
  tm_dots(col = "red", alpha = 0.7, size = 0.3) +
  tm_view(set.view = c(103.851959, 1.290270, 11)) +
  tm_layout(title = "Locations of New Massage")
```

# Fitness Centres

The spatial distribution of fitness centres is relatively extensive, stretching from the city centre to residential areas in both the east and west of Singapore, with a higher concentration in the eastern and central parts. This suggests that the fitness industry not only serves the central business districts but also penetrates into neighbourhoods, providing sports and leisure facilities for the community. The clusters around the Harbourfront area and new residential developments in the east also reflect the alignment of fitness consumption with emerging housing areas.

```{r}
tm_shape(biz_fitness_sf) +
  tm_dots(col = "green", alpha = 0.7, size = 0.3) +
  tm_view(set.view = c(103.851959, 1.290270, 11)) +
  tm_layout(title = "Locations of New Fitness Centres")
```

# Combine

This map shows the overall distribution of the three industries across Singapore. The points are mainly concentrated along major residential and commercial areas. In particular, the central and southern regions (such as the city centre and Orchard area) are densely clustered, indicating that clinics, massage parlours, and fitness centres tend to locate in core zones to serve areas with high population density and strong consumption power. Peripheral regions also have some distribution, but the points are noticeably sparser.

```{r}
biz_clinic_sf  <- biz_clinic_sf  %>% mutate(type = "Clinic")
biz_Massage_sf <- biz_Massage_sf %>% mutate(type = "Massage")
biz_fitness_sf <- biz_fitness_sf %>% mutate(type = "Fitness")

biz_all_sf <- rbind(biz_clinic_sf, biz_Massage_sf, biz_fitness_sf)

tmap_mode("view")

tm_shape(biz_all_sf) +
  tm_dots(col = "type", palette = c("blue","red","green"),
          alpha = 0.7, size = 0.2, legend.show = TRUE) +
  tm_view(set.view = c(103.851959, 1.290270, 11)) +
  tm_layout(title = "Locations of Clinics, Massage and Fitness Centres",
            legend.outside = TRUE)
```
:::

```{r}
tmap_mode("plot")
```

## 5.3 Singapore Basemap Overview

The map above presents the administrative subzone boundaries of Singapore as defined in the Urban Redevelopment Authority (URA) Master Plan 2019. Each polygon represents a planning subzone, which together form the basic spatial units for urban and statistical analysis. The map is projected in the SVY21 / Singapore TM coordinate system (EPSG:3414), which is the national standard for spatial data in Singapore.

This base map provides the essential geographic framework for subsequent spatial analyses. It allows for the overlay of business locations, demographic data, and other point or area-based information, enabling interpretation of spatial patterns across Singapore’s Central, East, West, North, and North-East regions. The inclusion of a scale bar and compass rose ensures that spatial orientation and measurement can be maintained in the analysis.

```{r}
tmap_style("natural")
tm_shape(mpsz) +
  tm_fill("lightgreen", title = "Singapore Boundary") +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Map of Singapore",
            main.title.position = "center",
            main.title.size = 1.0,
            legend.height = 0.35, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha = 0.2)
```

## 5.4 Quick Maps by Industry

These three graphs are visualizations of the distribution of data points, and they have not been combined with a map. From these points, it can be seen that the distribution of massage services is somewhat scattered in the non-central area, while the distributions of clinics and fitness centers are more concentrated, showing a clustered trend.

::: panel-tabset
# Clinics

```{r}
qtm(biz_clinic_sf)
```

# Massage

```{r}
qtm(biz_Massage_sf)
```

# Fitness Centres

```{r}
qtm(biz_fitness_sf)
```
:::

## 5.5 Regional Overlay Maps

::: panel-tabset
# Clinics

```{r}
tm_shape(mpsz) +
  tm_polygons("REGION_N", alpha = 0.3) +
tm_shape(biz_clinic_sf) +
  tm_dots(col = "blue", size = 0.05, popup.vars = c("entity_name", "ADDRESS"))
```

# Massage

```{r}
tm_shape(mpsz) +
  tm_polygons("REGION_N", alpha = 0.3) +
tm_shape(biz_Massage_sf) +
  tm_dots(col = "blue", size = 0.05, popup.vars = c("entity_name", "ADDRESS"))
```

# Fitness Centres

```{r}
tm_shape(mpsz) +
  tm_polygons("REGION_N", alpha = 0.3) +
tm_shape(biz_fitness_sf) +
  tm_dots(col = "blue", size = 0.05, popup.vars = c("entity_name", "ADDRESS"))
```
:::

## 5.6 Choropleth Maps by Subzone

```{r}
mpsz <- st_zm(mpsz, drop = TRUE)     
mpsz <- st_make_valid(mpsz)          
mpsz <- st_collection_extract(mpsz, "POLYGON")
mpsz <- st_cast(mpsz, "MULTIPOLYGON", warn = FALSE)
mpsz <- st_transform(mpsz, 3414)     
stopifnot(all(st_is_valid(mpsz)))
```

-   Fitness centres are moderately dispersed, balancing between city and neighbourhood coverage.

-   Massage services are highly centralised, clustered in commercial and entertainment hubs, with limited suburban presence.

-   Clinics are the most evenly spread and most numerous, with strong representation across nearly all regions, reflecting their role as a public necessity rather than a lifestyle or leisure option.

::: panel-tabset
# Clinics

Clinics exhibit the widest spatial coverage, appearing consistently across central, eastern, and suburban regions. Certain subzones, especially in the eastern part of the island, record over 20 or even 30 clinics, far exceeding the densities observed for fitness or massage services. This balanced yet intensive distribution pattern aligns with the essential service role of clinics, ensuring healthcare accessibility across both densely populated and peripheral residential areas.

```{r}
clinic_with_zone <- st_join(biz_clinic_sf, mpsz, join = st_within)

clinic_count <- clinic_with_zone %>%
  st_drop_geometry() %>%
  group_by(SUBZONE_N) %>%
  summarise(count = n())

mpsz_clinic <- mpsz %>%
  left_join(clinic_count, by = "SUBZONE_N")

tm_shape(mpsz_clinic) +
  tm_polygons("count", palette = "Blues", title = "Clinics") +
tm_shape(biz_clinic_sf) +
  tm_dots(col = "blue", size = 0.05)
```

# Massage

Massage businesses display a more uneven distribution, with denser clusters in the central and southern subzones (such as the city centre and Orchard area). Several subzones register 10 or more businesses, while many peripheral subzones have very few or none. This reflects the industry’s dependence on high-traffic commercial and leisure districts, highlighting its strong commercial orientation compared to fitness businesses.

```{r}
massage_with_zone <- st_join(biz_Massage_sf, mpsz, join = st_within)

massage_count <- massage_with_zone %>%
  st_drop_geometry() %>%
  group_by(SUBZONE_N) %>%
  summarise(count = n())

mpsz_massage <- mpsz %>%
  left_join(massage_count, by = "SUBZONE_N")

tm_shape(mpsz_massage) +
  tm_polygons("count", palette = "Blues", title = "Massage") +
tm_shape(biz_Massage_sf) +
  tm_dots(col = "red", size = 0.05)
```

# Fitness Centres

Fitness businesses are distributed across many subzones but generally show moderate clustering in the central and eastern regions. Most subzones only contain 1 to 5 businesses, with a few areas reaching 10 or more. The spatial pattern suggests that fitness centres are present in both residential neighbourhoods and near city-fringe areas, aiming to serve everyday community needs rather than concentrating solely in the central business district.

```{r}
fitness_with_zone <- st_join(biz_fitness_sf, mpsz, join = st_within)

fitness_count <- fitness_with_zone %>%
  st_drop_geometry() %>%
  group_by(SUBZONE_N) %>%
  summarise(count = n())

mpsz_fitness <- mpsz %>%
  left_join(fitness_count, by = "SUBZONE_N")

tm_shape(mpsz_fitness) +
  tm_polygons("count", palette = "Blues", title = "Fitness Businesses") +
tm_shape(biz_fitness_sf) +
  tm_dots(col = "red", size = 0.05)
```
:::

# 6 Point Pattern Preparation

## 6.1 Converting Boundary to Owin Object

In this step, the Master Plan 2019 Subzone Boundary (mpsz) dataset, which was originally stored as an sf object, was converted into an owin (observation window) object using the as.owin() function. The owin format is a fundamental spatial data structure in the spatstat package, required for point pattern analysis as it defines the study region (the spatial window within which points are observed and analyzed).

After conversion, we checked the class of the object to ensure it was successfully transformed, and then plotted the boundary. From the figure below, we can see the resulting map boundary of Singapore, which will serve as the analytical window for subsequent spatial point pattern analysis.

```{r}
sg_owin <- as.owin(mpsz)
```

```{r}
class(sg_owin)
```

```{r}
plot(sg_owin)
```

## 6.2 Converting Business Data to PPP Objects

In this step, the three business datasets (clinics, massage services, and fitness centres), which were initially stored as sf objects, were converted into ppp (planar point pattern) objects using the as.ppp() function from the spatstat package. The ppp format is essential for spatial point pattern analysis, as it represents each business location as a point within a defined observation window.

::: panel-tabset
# Clinics

```{r}
clinic_ppp <- as.ppp(biz_clinic_sf)
class(clinic_ppp)
```

# Massage

```{r}
Massage_ppp <- as.ppp(biz_Massage_sf)
class(Massage_ppp)
```

# Fitness Centres

```{r}
fitness_ppp <- as.ppp(biz_fitness_sf)
class(fitness_ppp)
```
:::

## 6.3 Cropping Point Patterns to Study Window

After converting the business datasets into ppp objects, each point pattern was cropped to the Singapore study window (sg_owin). This ensures that only the points located within the official Singapore boundary are retained for analysis, while any points falling outside the boundary are removed.

We can see the corresponding data quantity for each item from the output results below.

::: panel-tabset
# Clinics

```{r}
clinic_ppp = clinic_ppp[sg_owin]
```

```{r}
clinic_ppp
```

# Massage

```{r}
Massage_ppp = Massage_ppp[sg_owin]
```

```{r}
Massage_ppp
```

# Fitness Centres

```{r}
fit_ppp = fitness_ppp[sg_owin]
```

```{r}
fit_ppp
```
:::

# 7 First-Order Spatial Analysis

## 7.1 Nearest Neighbour Analysis (NNA)

Nearest Neighbor Analysis (NNA) is a spatial statistics method that calculates the average distance between each point and its closest neighbor to determine if a pattern of points is clustered, dispersed, or randomly distributed.

Clark-Evans test is a specific statistical method used within NNA to quantify whether a point pattern is clustered, random, or uniformly spaced, using the Clark-Evans aggregation index (R) to describe this pattern. NNA provides a numerical value that describes the degree of clustering or regularity, and the Clark-Evans test calculates a specific index (R) for this purpose.

::: panel-tabset
# Clinics

• H₀ (Null Hypothesis): The distribution of clinics is randomly distributed. • H₁ (Alternative Hypothesis): The distribution of clinics is not randomly distributed.

Result: R = 0.38104, p-value \< 2.2e-16. Reject H₀. The distribution of clinics is not random, with a clustered pattern across Singapore.

```{r}
clarkevans.test(clinic_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"))
```

# Massage

• H₀ (Null Hypothesis): The distribution of massage services is randomly distributed. • H₁ (Alternative Hypothesis): The distribution of massage services is not randomly distributed.

Result: R = 0.33504, p-value \< 2.2e-16. Reject H₀. The distribution of massage services is not random, showing the strongest clustering among the three industries.

```{r}
clarkevans.test(Massage_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"))
```

# Fitness Centres

• H₀ (Null Hypothesis): The distribution of fitness centres is randomly distributed. • H₁ (Alternative Hypothesis): The distribution of fitness centres is not randomly distributed.

Result: R = 0.39117, p-value \< 2.2e-16. Reject H₀. The distribution of fitness centres is not random, but significantly clustered.

```{r}
clarkevans.test(fit_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"))
```
:::

## 7.2 Kernel Density Estimation (KDE)

Kernel Density Estimation (KDE) is a valuable tool for visualising and analyzing first-order spatial point patterns. It is widely considered a method within Exploratory Spatial Data Analysis (ESDA) because it’s used to visualize and understand spatial data patterns by transforms discrete point data (like locations of childcare service, crime incidents or disease cases) into continuous density surfaces that reveal clusters and variations in event occurrences, without making prior assumptions about data distribution. It helps to begin understanding data distribution, identify hotspots, and explore relationships between spatial variables before performing more rigorous analysis.

### 7.2.1 Automatic bandwidth selection method

::: panel-tabset
# Clinics

```{r}
clinic_ppp_km <- rescale.ppp(
  clinic_ppp, 1000, "km")

kde_clinic_diggle_km <- density(clinic_ppp_km,
                                sigma = bw.diggle,
                                edge = TRUE,
                                kernel = "gaussian")

plot(kde_clinic_diggle_km, main = "Clinic KDE (bw.diggle, km)")
```

# Massage

```{r}
massage_ppp_km <- rescale.ppp(
  Massage_ppp, 1000, "km")

kde_massage_diggle_km <- density(massage_ppp_km,
                                 sigma = bw.diggle,
                                 edge = TRUE,
                                 kernel = "gaussian")

plot(kde_massage_diggle_km, main = "Massage KDE (bw.diggle, km)")
```

# Fitness Centres

```{r}
fit_ppp_km <- rescale.ppp(
  fit_ppp, 1000, "km")

kde_fit_diggle_km <- density(fit_ppp_km,
                              sigma=bw.diggle,
                              edge=TRUE,
                              kernel="gaussian")
plot(kde_fit_diggle_km)
```
:::

### 7.2.2 Adaptive bandwidth

::: panel-tabset
# Clinics

```{r}
kde_clinic_ab <- adaptive.density(
  clinic_ppp_km, 
  method="kernel")
plot(kde_clinic_ab)
```

# Massage

```{r}
kde_massage_ab <- adaptive.density(
  massage_ppp_km, 
  method="kernel")
plot(kde_massage_ab)
```

# Fitness Centres

```{r}
kde_fit_ab <- adaptive.density(
  fit_ppp_km, 
  method="kernel")
plot(kde_fit_ab)
```
:::

### 7.2.3 Cartographic quality KDE map

```{r}
kde_clinic_bw_terra <- rast(kde_clinic_diggle_km)
crs(kde_clinic_bw_terra) <- "EPSG:3414"
class(kde_clinic_bw_terra)
kde_clinic_bw_terra

kde_massage_bw_terra <- rast(kde_massage_diggle_km)
crs(kde_massage_bw_terra) <- "EPSG:3414"
class(kde_massage_bw_terra)
kde_massage_bw_terra

kde_fit_bw_terra <- rast(kde_fit_diggle_km)
crs(kde_fit_bw_terra) <- "EPSG:3414"
class(kde_fit_bw_terra)
kde_fit_bw_terra
```

::: panel-tabset
# Clinics

```{r}
tm_shape(kde_clinic_bw_terra) + 
  tm_raster(col.scale = 
              tm_scale_continuous(
                values = "viridis"),
            col.legend = tm_legend(
            title = "Density values",
            title.size = 0.7,
            text.size = 0.7,
            bg.color = "white",
            bg.alpha = 0.7,
            position = tm_pos_in(
              "right", "bottom"),
            frame = TRUE)) +
  tm_graticules(labels.size = 0.7) +
  tm_compass() +
  tm_layout(scale = 1.0)
```

# Massage

```{r}
tm_shape(kde_massage_bw_terra) + 
  tm_raster(col.scale = 
              tm_scale_continuous(
                values = "viridis"),
            col.legend = tm_legend(
            title = "Density values",
            title.size = 0.7,
            text.size = 0.7,
            bg.color = "white",
            bg.alpha = 0.7,
            position = tm_pos_in(
              "right", "bottom"),
            frame = TRUE)) +
  tm_graticules(labels.size = 0.7) +
  tm_compass() +
  tm_layout(scale = 1.0)
```

# Fitness Centres

```{r}
tm_shape(kde_fit_bw_terra) + 
  tm_raster(col.scale = 
              tm_scale_continuous(
                values = "viridis"),
            col.legend = tm_legend(
            title = "Density values",
            title.size = 0.7,
            text.size = 0.7,
            bg.color = "white",
            bg.alpha = 0.7,
            position = tm_pos_in(
              "right", "bottom"),
            frame = TRUE)) +
  tm_graticules(labels.size = 0.7) +
  tm_compass() +
  tm_layout(scale = 1.0)
```
:::

# 8 Second-Order Spatial Analysis

## 8.1 Subset by Planning area

From the mpsz dataset, five sub-datasets, namely Jurong East/West, Newton, Serangoon, and Woodlands, were selected based on the planning areas (PLN_AREA_N) for facilitating the analysis by regions.

```{r}
je <- mpsz %>%
  filter(PLN_AREA_N == "JURONG EAST")
jw <- mpsz %>%
  filter(PLN_AREA_N == "JURONG WEST")
nt <- mpsz %>%
  filter(PLN_AREA_N == "NEWTON")
se <- mpsz %>%
  filter(PLN_AREA_N == "SERANGOON")
wl <- mpsz %>%
  filter(PLN_AREA_N == "WOODLANDS")
```

## 8.2 Plot Planning Area Geometries

Quickly visualize the polygon geometries of the five planning areas, check whether the boundaries are correct, and whether there are any invalid geometries or coordinate issues.

```{r}
par(mfrow=c(2,2))
plot(st_geometry(je), main = "JURONG EAST")
plot(st_geometry(jw), main = "JURONG WEST")
plot(st_geometry(nt), main = "NEWTON")
plot(st_geometry(se), main = "SERANGOON")
plot(st_geometry(wl), main = "WOODLANDS")
```

## 8.3 Convert Planning Areas to Windows (owin)

Convert each planning area polygon into an observation window owin in spatstat, which will serve as the research boundary for subsequent point pattern clipping and analysis.

```{r}
je_owin = as.owin(je)
jw_owin = as.owin(jw)
nt_owin = as.owin(nt)
se_owin = as.owin(se)
wl_owin = as.owin(wl)
```

## 8.4 Point Patterns per Planning Area (Clinics / Massage / Fitness)

The point pattern data were clipped to their corresponding planning area windows to retain only events within each boundary. The patterns were then rescaled from meters to kilometers using rescale.ppp(..., 1000, "km"), and visualized with plot(unmark(...)) to provide a quick view of the distribution, allowing an intuitive comparison of density and spatial structure across regions.

::: panel-tabset
# Clinics

```{r}
clinic_je_ppp  = clinic_ppp[je_owin]
clinic_jw_ppp  = clinic_ppp[jw_owin]
clinic_nt_ppp  = clinic_ppp[nt_owin]
clinic_se_ppp  = clinic_ppp[se_owin]
clinic_wl_ppp  = clinic_ppp[wl_owin]

clinic_je_ppp.km  = rescale.ppp(clinic_je_ppp, 1000, "km")
clinic_jw_ppp.km  = rescale.ppp(clinic_jw_ppp, 1000, "km")
clinic_nt_ppp.km  = rescale.ppp(clinic_nt_ppp, 1000, "km")
clinic_se_ppp.km  = rescale.ppp(clinic_se_ppp, 1000, "km")
clinic_wl_ppp.km  = rescale.ppp(clinic_wl_ppp, 1000, "km")

par(mfrow = c(2, 3), cex = 0.3)
plot(unmark(clinic_je_ppp.km), main = "JURONG EAST (Clinic)")
plot(unmark(clinic_jw_ppp.km), main = "JURONG WEST (Clinic)")
plot(unmark(clinic_nt_ppp.km), main = "NEWTON (Clinic)")
plot(unmark(clinic_se_ppp.km), main = "SERANGOON (Clinic)")
plot(unmark(clinic_wl_ppp.km), main = "WOODLANDS (Clinic)")
```

# Massage

```{r}
massage_je_ppp  = Massage_ppp[je_owin]
massage_jw_ppp  = Massage_ppp[jw_owin]
massage_nt_ppp  = Massage_ppp[nt_owin]
massage_se_ppp  = Massage_ppp[se_owin]
massage_wl_ppp  = Massage_ppp[wl_owin]

massage_je_ppp.km  = rescale.ppp(massage_je_ppp, 1000, "km")
massage_jw_ppp.km  = rescale.ppp(massage_jw_ppp, 1000, "km")
massage_nt_ppp.km  = rescale.ppp(massage_nt_ppp, 1000, "km")
massage_se_ppp.km  = rescale.ppp(massage_se_ppp, 1000, "km")
massage_wl_ppp.km  = rescale.ppp(massage_wl_ppp, 1000, "km")

par(mfrow = c(2, 3), cex = 0.3)
plot(unmark(massage_je_ppp.km), main = "JURONG EAST (Massage)")
plot(unmark(massage_jw_ppp.km), main = "JURONG WEST (Massage)")
plot(unmark(massage_nt_ppp.km), main = "NEWTON (Massage)")
plot(unmark(massage_se_ppp.km), main = "SERANGOON (Massage)")
plot(unmark(massage_wl_ppp.km), main = "WOODLANDS (Massage)")
```

# Fitness Centres

```{r}
fit_je_ppp  = fitness_ppp[je_owin]
fit_jw_ppp  = fitness_ppp[jw_owin]
fit_nt_ppp  = fitness_ppp[nt_owin]
fit_se_ppp  = fitness_ppp[se_owin]
fit_wl_ppp  = fitness_ppp[wl_owin]

fit_je_ppp.km  = rescale.ppp(fit_je_ppp, 1000, "km")
fit_jw_ppp.km  = rescale.ppp(fit_jw_ppp, 1000, "km")
fit_nt_ppp.km  = rescale.ppp(fit_nt_ppp, 1000, "km")
fit_se_ppp.km  = rescale.ppp(fit_se_ppp, 1000, "km")
fit_wl_ppp.km  = rescale.ppp(fit_wl_ppp, 1000, "km")

par(mfrow = c(2, 3), cex = 0.3)
plot(unmark(fit_je_ppp.km), main = "JURONG EAST (Fitness)")
plot(unmark(fit_jw_ppp.km), main = "JURONG WEST (Fitness)")
plot(unmark(fit_nt_ppp.km), main = "NEWTON (Fitness)")
plot(unmark(fit_se_ppp.km), main = "SERANGOON (Fitness)")
plot(unmark(fit_wl_ppp.km), main = "WOODLANDS (Fitness)")
```
:::

## 8.5 Setting Random Seed

The random seed is fixed with set.seed(1234) to ensure reproducibility of stochastic spatial simulations and statistical tests, so that the results can be replicated consistently.

```{r}
set.seed(1234)
```

## 8.6 Second-Order Spatial Analysis For Clinic

Ho = The distribution of Clinic at SERANGOON are randomly distributed.

H1= The distribution of Clinic at SERANGOON are not randomly distributed.

Taken together, the K and L functions—and to a lesser extent the G function—show significant departures above the CSR envelopes over specific distance bands, indicating scale-dependent clustering of clinics in Serangoon. Although the F function does not reject CSR, the convergent evidence from K and L (which are more sensitive to cumulative interaction over distance) supports the conclusion that the pattern is not fully random. Overall, we reject H₀ and infer that clinic locations in Serangoon exhibit statistically significant clustering at particular spatial scales, while acknowledging that the effect is not uniform across all distances.

::: panel-tabset
# G-function

Ho = The distribution of Clinic at SERANGOON are randomly distributed.

H1= The distribution of Clinic at SERANGOON are not randomly distributed.

The observed curve exceeds the CSR envelope over parts of the distance range, indicating statistically significant clustering at those scales; elsewhere it remains within the envelope. Decision: Reject H₀ only for the distances where the curve is above the envelope (scale-dependent effect).

```{r}
G_clinic_se = Gest(clinic_se_ppp, correction = "best")
plot(G_clinic_se, xlim=c(0,500))
```

```{r}

G_clinic_se.csr <- envelope(clinic_se_ppp, Gest, nsim = 999)
```

```{r}
plot(G_clinic_se.csr)
```

# F-function

Ho = The distribution of Clinic at SERANGOON are randomly distributed.

H1= The distribution of Clinic at SERANGOON are not randomly distributed.

The observed curve lies entirely within the CSR envelope, providing no statistically significant departure from CSR. Decision: Fail to reject H₀.

```{r}
F_clinic_se = Fest(clinic_se_ppp, correction = "best")
plot(F_clinic_se)
```

```{r}

F_clinic_se.csr <- envelope(clinic_se_ppp, Fest, nsim = 999)
```

```{r}
plot(F_clinic_se.csr)
```

# K-function

Ho = The distribution of Clinic at SERANGOON are randomly distributed.

H1= The distribution of Clinic at SERANGOON are not randomly distributed.

The observed curve rises above the upper CSR envelope across portions of the range, implying more neighbours than expected under CSR. Decision: Reject H₀ for those distance ranges, consistent with clustering.

```{r}
K_clinic_se = Kest(clinic_se_ppp, correction = "Ripley")
plot(K_clinic_se, . -r ~ r, ylab= "K(d)-r", xlab = "d(m)")
```

```{r}
K_clinic_se.csr <- envelope(clinic_se_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(K_clinic_se.csr, . - r ~ r, xlab="d", ylab="K(d)-r")
```

# L-function

Ho = The distribution of Clinic at SERANGOON are randomly distributed.

H1= The distribution of Clinic at SERANGOON are not randomly distributed.

The variance-stabilised curve similarly exceeds the CSR envelope over certain distances, corroborating the clustering signal. Decision: Reject H₀ for the distances where the curve is above the envelope.

```{r}
L_clinic_se = Lest(clinic_se_ppp, correction = "Ripley")
plot(L_clinic_se, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)")
```

```{r}
#| cache: true
L_clinic_se.csr <- envelope(clinic_se_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(L_clinic_se.csr, . - r ~ r, xlab="d", ylab="L(d)-r")
```
:::

## 8.7 Second-Order Spatial Analysis For Massage

Ho = The distribution of Massage at SERANGOON are randomly distributed.

H1= The distribution of Massage at SERANGOON are not randomly distributed.

Although the F-function does not reject CSR and the G-function shows only slight departures at limited scales, the K and L functions provide robust evidence of clustering across a broad range of distances, with L indicating consistent, scale-spanning departures from CSR. Given that K and L integrate information over neighborhoods and are more sensitive to cumulative interaction than G and F, the weight of evidence supports rejection of H₀: massage establishments in Serangoon are not randomly distributed but exhibit statistically significant clustering, with the effect strongest and most pervasive at medium to larger spatial scales.

::: panel-tabset
# G-function

Ho = The distribution of Massage at SERANGOON are randomly distributed.

H1= The distribution of Massage at SERANGOON are not randomly distributed.

The observed curve only marginally exceeds the CSR envelope over a very limited distance range, indicating weak, scale-limited clustering; elsewhere it remains within the envelope. Decision: Reject H₀ only at those narrow distances; otherwise fail to reject.

```{r}
G_massage_se = Gest(massage_se_ppp, correction = "best")
plot(G_massage_se, xlim=c(0,500))
```

```{r}
G_massage_se.csr <- envelope(massage_se_ppp, Gest, nsim = 999)
```

```{r}
plot(G_massage_se.csr)
```

# F-function

Ho = The distribution of Massage at SERANGOON are randomly distributed.

H1= The distribution of Massage at SERANGOON are not randomly distributed.

The observed curve lies entirely within the CSR envelope, providing no statistically significant departure from CSR. Decision: Fail to reject H₀.

```{r}
F_massage_se = Fest(massage_se_ppp, correction = "best")
plot(F_massage_se)
```

```{r}
F_massage_se.csr <- envelope(massage_se_ppp, Fest, nsim = 999)
```

```{r}
plot(F_massage_se.csr)
```

# K-function

Ho = The distribution of Massage at SERANGOON are randomly distributed.

H1= The distribution of Massage at SERANGOON are not randomly distributed.

The observed curve is above the CSR envelope for most distances, dipping inside only briefly, implying more neighbors than expected under CSR across a wide range of scales. Decision: Reject H₀ (clustering) over most distances.

```{r}
K_massage_se = Kest(massage_se_ppp, correction = "Ripley")
plot(K_massage_se, . -r ~ r, ylab= "K(d)-r", xlab = "d(m)")
```

```{r}
K_massage_se.csr <- envelope(massage_se_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(K_massage_se.csr, . - r ~ r, xlab="d", ylab="K(d)-r")
```

# L-function

Ho = The distribution of Massage at SERANGOON are randomly distributed.

H1= The distribution of Massage at SERANGOON are not randomly distributed.

The variance-stabilized curve remains above the CSR envelope across all distances considered, offering strong and consistent evidence of clustering. Decision: Reject H₀.

```{r}
L_massage_se = Lest(massage_se_ppp, correction = "Ripley")
plot(L_massage_se, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)")
```

```{r}
L_massage_se.csr <- envelope(massage_se_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(L_massage_se.csr, . - r ~ r, xlab="d", ylab="L(d)-r")
```
:::

## 8.8 Second-Order Spatial Analysis For Fitness

Ho = The distribution of Fitness at SERANGOON are randomly distributed.

H1= The distribution of Fitness at SERANGOON are not randomly distributed.

Taken together, G shows clustering confined to very short distances (0–100 m), while K and L demonstrate statistically significant clustering up to \~600 m, beyond which both functions revert to compatibility with CSR. The F-function does not reject CSR at any scale, which is not uncommon given its different sensitivity (distance from random locations). Overall, the convergent evidence from K and L—supported by G at the smallest scales—indicates that fitness establishments in Serangoon are not randomly distributed, but exhibit scale-dependent clustering concentrated at short to medium ranges (≤ 600 m); at larger scales, the pattern is indistinguishable from CSR. Accordingly, we reject H₀ for distances below 600 m and fail to reject H₀ at greater distances.

::: panel-tabset
# G-function

Ho = The distribution of Fitness at SERANGOON are randomly distributed.

H1= The distribution of Fitness at SERANGOON are not randomly distributed.

The observed curve exceeds the CSR envelope only at 0–100 m, indicating short-range clustering; elsewhere it falls within the envelope. Decision: Reject H₀ at 0–100 m only; otherwise fail to reject.

```{r}
G_fit_se = Gest(fit_se_ppp, correction = "best")
plot(G_fit_se, xlim=c(0,500))
```

```{r}
#| cache: true
G_fit_se.csr <- envelope(fit_se_ppp, Gest, nsim = 999)
```

```{r}
plot(G_fit_se.csr)
```

# F-function

Ho = The distribution of Fitness at SERANGOON are randomly distributed.

H1= The distribution of Fitness at SERANGOON are not randomly distributed.

The observed curve remains entirely within the CSR envelope across all distances, providing no significant departure from CSR. Decision: Fail to reject H₀.

```{r}
F_fit_se = Fest(fit_se_ppp, correction = "best")
plot(F_fit_se)
```

```{r}
F_fit_se.csr <- envelope(fit_se_ppp, Fest, nsim = 999)
```

```{r}
plot(F_fit_se.csr)
```

# K-function

Ho = The distribution of Fitness at SERANGOON are randomly distributed.

H1= The distribution of Fitness at SERANGOON are not randomly distributed.

The observed curve lies above the CSR envelope for r \< 600 m and falls within it for r ≥ 600 m, implying scale-limited clustering concentrated at short to medium ranges. Decision: Reject H₀ for r \< 600 m; fail to reject for r ≥ 600 m.

```{r}
K_fit_se = Kest(fit_se_ppp, correction = "Ripley")
plot(K_fit_se, . -r ~ r, ylab= "K(d)-r", xlab = "d(m)")
```

```{r}
K_fit_se.csr <- envelope(fit_se_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(K_fit_se.csr, . - r ~ r, xlab="d", ylab="K(d)-r")
```

# L-function

Ho = The distribution of Fitness at SERANGOON are randomly distributed.

H1= The distribution of Fitness at SERANGOON are not randomly distributed.

The variance-stabilized curve is above the CSR envelope for r \< 600 m and inside the envelope for r ≥ 600 m, corroborating short- to mid-range clustering with no large-scale departure. Decision: Reject H₀ for r \< 600 m; fail to reject for r ≥ 600 m.

```{r}
L_fit_se = Lest(fit_se_ppp, correction = "Ripley")
plot(L_fit_se, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)")
```

```{r}
#| cache: true
L_fit_se.csr <- envelope(fit_se_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(L_fit_se.csr, . - r ~ r, xlab="d", ylab="L(d)-r")
```
:::

# 9 Spatio-Temporal Point Patterns Analysis

## 9.1 Data Prepare for Spatio-Temporal Point Patterns Analysis

This code prepares the business datasets for spatio-temporal analysis in the Serangoon planning area. First, the study period is defined from January 2024 to June 2025, and a complete sequence of months within this range is generated and formatted as categorical labels (e.g., Jan 2024), which serve as ordered factors for subsequent faceted visualizations. The Serangoon boundary polygon is then extracted from the URA Master Plan 2019 subzone dataset to delimit the study region.

A custom function, prep_points(), is implemented to standardize data preprocessing across all three industry types (clinics, massage services, and fitness centres). The function clips point locations to the Serangoon boundary, filters records by the defined time range, and assigns the corresponding YearMonth factor to each entry. Applying this function produces three cleaned subsets: biz_clinic_se, biz_massage_se, and biz_fitness_se.

Finally, a facetable polygon layer (mpsz_se_facets) is constructed by replicating the Serangoon boundary for each month in the study period and attaching the appropriate YearMonth label. This ensures that monthly facet maps will display consistently, even for months with no business registrations.

Overall, this procedure integrates both spatial and temporal dimensions, generating harmonized datasets and a reusable polygon base layer to support monthly faceted mapping and subsequent spatial analyses.

```{r}
range_start <- as.Date("2024-01-01")
range_end   <- as.Date("2025-06-30")

all_months <- seq(as.Date("2024-01-01"), as.Date("2025-06-01"), by = "1 month")
all_levels <- format(all_months, "%b %Y")

mpsz_se <- mpsz %>% dplyr::filter(PLN_AREA_N == "SERANGOON")
stopifnot(nrow(mpsz_se) > 0)

prep_points <- function(sf_points, boundary_sf, start_date, end_date, levels_vec){
  sf::st_join(sf_points, boundary_sf %>% dplyr::select(PLN_AREA_N),
              join = sf::st_within, left = FALSE) %>%
    dplyr::filter(date >= start_date, date <= end_date) %>%
    dplyr::mutate(YearMonth = factor(format(date, "%b %Y"),
                                     levels = levels_vec, ordered = TRUE))
}

biz_fitness_se <- prep_points(biz_fitness_sf,  mpsz_se, range_start, range_end, all_levels)
biz_clinic_se  <- prep_points(biz_clinic_sf,   mpsz_se, range_start, range_end, all_levels)
biz_massage_se <- prep_points(biz_Massage_sf,  mpsz_se, range_start, range_end, all_levels)

mpsz_se_facets <- dplyr::bind_rows(lapply(all_levels, function(m) {
  tmp <- mpsz_se %>% dplyr::select(geometry)
  tmp$YearMonth <- factor(m, levels = all_levels, ordered = TRUE)
  tmp
})) %>% sf::st_as_sf()
```

## 9.2 Visuaising geographic distribution of 3 industry by month

Both clinics and fitness centres demonstrate broad community coverage, indicating that they serve everyday living needs (healthcare + wellness). Massage services, however, follow a different logic, being more dependent on commercial and leisure demand, and showing lower overlap with the other two. Overall, the three industries exhibit complementarity rather than competition: clinics ensure essential healthcare, fitness centres promote healthy lifestyles, and massage services satisfy scattered leisure needs.

::: panel-tabset
# Clinic

Clinics in Serangoon show the widest and most stable spatial distribution, covering different subzones. This reflects the public service nature of clinics, which are designed to meet residents’ healthcare needs across the area. Between February and June 2024 there were no new establishments, but from March to June 2025 multiple new clinics appeared each month, which may be related to policy changes.

```{r}
tm_shape(mpsz_se_facets, is.master = TRUE) +
  tm_polygons(fill = "grey90") +
  tm_borders(col = "grey60") +
tm_shape(biz_clinic_se) +
  tm_dots(size = 0.3, col = "blue") +
tm_facets(
  by = "YearMonth",
  ncol = 6, nrow = 3,
  drop.units = TRUE,       
  drop.empty.facets = FALSE, 
  free.coords = FALSE     
) +
tm_title("Clinics in SERANGOON by Month")

```

# Massage

Massage services are the least numerous and most spatially limited among the three industries. Over the 18-month period, only a handful of new outlets were registered, with most months recording just one. Their distribution is highly scattered, often located in central or easily accessible parts of Serangoon. This pattern reflects the commercial and leisure-oriented nature of massage services, which differ from clinics in their lack of uniform distribution. Compared with clinics and fitness centres, massage services have a weaker presence in Serangoon and represent a niche industry.

```{r}
tm_shape(mpsz_se_facets, is.master = TRUE) +
  tm_polygons(fill = "grey90") +
  tm_borders(col = "grey60") +
tm_shape(biz_massage_se) +
  tm_dots(size = 0.3, col = "red") +
tm_facets(
  by = "YearMonth",
  ncol = 6, nrow = 3,
  drop.units = TRUE,
  drop.empty.facets = FALSE,
  free.coords = FALSE
) +
tm_title("Massage  in SERANGOON by Month")
```

# Fitness

Fitness centres are moderate in scale but unevenly distributed. In some months multiple new establishments were registered, while in others none appeared. Their locations are widely spread, covering both central and peripheral subzones, with some small-scale clusters observed. Compared with clinics, fitness centres display greater volatility in registration, reflecting market fluctuations, shifting consumer demand, and industry competition.

```{r}
tm_shape(mpsz_se_facets, is.master = TRUE) +
  tm_polygons(fill = "grey90") +
  tm_borders(col = "grey60") +
tm_shape(biz_fitness_se) +
  tm_dots(size = 0.3) +
tm_facets(
  by = "YearMonth",
  ncol = 6, nrow = 3,
  drop.units = TRUE,       
  drop.empty.facets = FALSE,
  free.coords = FALSE       
)+
tm_title("Fitness centres  in SERANGOON by Month") 
```
:::


## 9.3 Spatio-temporal Point Patterns Analysis: stpp methods

```{r}
coords <- st_coordinates(biz_clinic_sf)
```

```{r}
clinic_df <- data.frame(
  x = coords[, 1],  
  y = coords[, 2],
  t = biz_clinic_sf$`date`)

clinic_stpp <- as.3dpoints(clinic_df)

class(clinic_stpp)
```

```{r}
plot(clinic_stpp)
```
```{r}
kbb_clinic <- STIKhat(clinic_stpp)
```

```{r}
plotK(kbb_clinic)
```

# 10 Reference

-   Kam, T. S. (2024). *R for Geospatial Data Science and Analytics* <https://r4gdsa.netlify.app>
